### 2015-11-07 paper.tex
Inital thoughts on the tpcp projet

\documentclass[11pt]{article}

\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{url}
\usepackage[english]{babel}
\usepackage{vmargin}
\usepackage{times}
\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
\usepackage{cite}
\usepackage{titling}
\usepackage{color}

\widowpenalty=10000
\clubpenalty=10000

\begin{document}

\title{Total Platform Cyber Protection}
\date{}

\maketitle

\section{Introduction}

Traditional security defenses for the computing environment have
focused on securing the \emph{boarder} of the network, as this is the
entry place of most attacks. Specifically, technologies such as
firewalls, intrusion detection systems, and automated threat forensics
operate at the network gateway, attempting to effectively secure the
boarder of the network.

However, even with such technologies in place, there is a rash of
security breach and compromises. Target. OPM. Trump Industries.

The goal of this work is to focus research effort on \emph{securing
  the entire computing platform.} An attack must effectively target a
specific vulnerability in a specific layer of the computing stack, and
an attacker uses that vulnerability to establish persistence on the
machine, potentially attacking the underlying computing layers of the
same machine or attacking other machines. Therefore, if we wish to
increase the security of our computing systems, and reduce the number
and scope of security breaches, it is essential that we encourage
focus on novel ideas, algorithms, and techniques to secure every level
of the computing stack.

In this paper, we will first discuss the computing stack, and, for
every layer, we will identify focus areas that attempt to secure that
layer. For each of those focus areas we will answer the following
questions:

\begin{itemize}
  \item What's the technical challenge for the focus area?
  \item Why is it hard?
  \item What are the open problems?
  \item How is the area addressed in the past (from a cyber perspective)?
  \item Are there intellectual tools that need to be brought to bear on the problem?
  \item How/when do we achieve results?
  \item Near-term and the long-term?
\end{itemize}

\section{Background}

What is the background of the computing layers? Here, having the
layers in a graphic, explaining the interconnected layers would be
good.

\section{Hardware}

Important problems include how to deal with counterfeit chips, better
non-destructive examination methods, continuous monitoring of the
integrity not just at boot time, and detection of backdoors and 
``undocumented features''.  Hardware design can also be done to
aid this process, e.g., having some part of the chip consume more power
to make it more identifiable for authentication purposes.  Other
possibilities include running multiple chips with different
implementations side-by-side and comparing results to detect unexpected
behavior.

Identification and signatures using features that are more robust to
aging/use over time

\subsection{Focus Areas}

\begin{enumerate}
	\item Non-destructive IC examination, identification, and
	authentication
	\item Continuous or near-continuous physical monitoring for
	establishing root of trust
	\item Inspection-friendly design for hardware transparency
\end{enumerate}

\section{Firmware}

Firmware must deal with the fact that no OS-style abstraction layer
is available, so hardware interaction is important.  Reverse
engineering is an interesting at the firmware level because most
vendors want to make it difficult to protect IP; it is also more
tedious and would the RE process would benefit more from better tools.

\subsection{Focus Areas}

\begin{enumerate}
	\item Decouple firmware from the hardware (emulation) to aid analysis
	\item Better tools for automated reverse engineering, including
	analysis of layouts, modules, determining code vs. data
	\item How to patch firmware with analysis and guarantees about
	system behavior
	\begin{enumerate}
		\item did you fix the vulnerabilities?
		\item did you introduce new vulnerabilities?
		\item did your changes disrupt the timing of the original firmware?
	\end{enumerate}
\end{enumerate}

\section{Bus and Interconnect}

\subsection{Focus Areas}

Critical tech challenges:
\begin{enumerate}
	\item Lightweight device authentication that is bus-appropriate
	\item Operational mode setting to determine allowed capabilities
	for each device
\end{enumerate}

\section{Hypervisor}

\subsection{Focus Areas}

Critical tech challenges:
\begin{enumerate}
	\item Protect private information within a VM/guest from spying
	and interception by the host, bootstrap image with private key
	\item Attest memory, e.g., did my cloud provider put something in
	to read memory?
	\item Achieve predictable real-time scheduling in a lightweight
	and secure hypervisor
\end{enumerate}

\section{Operating Systems}

Discussion centered on reducing the attack surface by stripping out
unnecessary and unneeded functionalities.

\subsection{Focus Areas}

\begin{enumerate}
	\item Automated reduction and pruning of features to a minimal set
	of required functionality
	\item Non-control data flow attacks
	\item Concurrency-based attacks
\end{enumerate}

\section{Middleware / Runtime}

This focus of this layer includes the runtime environments and abstractions
below a high-level programming language that convert it into bytecode.

\subsection{Focus Areas}

\begin{enumerate}
	\item Dealing with bloat and reducing the attack surface
	\item Instrumentation is easier at this layer, but faster ways to observe execution are needed
	\item Controlling easily misued middleware APIs (confused deputy)
\end{enumerate}

\section{Application Layer}

Much current work focuses on just looking at memory corruption bugs, but
we need to look at how the same sort of advanced protections can be
applied to other classes of bugs, i.e., what is next?  This includes:
input validation, logic problems, leaking secrets, performance degradation,
denial-of-service, command (e.g., SQL) injection, and concurrency attacks.

\subsection{Focus Areas}

\begin{enumerate}
	\item The next generation of vulnerabilities (e.g., logic flaws)
	\item Need good tools for analysis of race conditions / concurrency attacks
\end{enumerate}

\section{Network Layer}

\subsection{Focus Areas}

Discussion centered on new / unexplored ways in which to secure networking
protocols.  One primary area that was discussed centered around
applying artificial diversity to protocols,  with implementation
including potentially taking advantage of the ambiguity of protocols
(i.e., differences in how messages in different protocols are
interpreted by endpoints and middleboxes).  Some group members also
felt like static and dynamic security analysis (to include fuzzing) of
network protocols was an area that needed more work.

\begin{enumerate}

  \item Protocol diversity for surviving attacks

  \item Raise the bar on an attacker by forcing them to attack all
  channels to be successful

  \item Protocol dissection is not a solved problem (i.e., how to
  understand and analyze new and unknown protocols)

  \begin{enumerate}
  	\item Profile traffic to application-specific communication for
  	composing into system-specific configuration baseline
  \end{enumerate}

\end{enumerate}

\section{User Layer}

Discussion centered on trust of sensors in a system.  For instance,
what can be done about sensors that lie (e.g., because they have been
compromised), and also how to force a hacker to own all the sensors in
a system in order to truly take control.  Control theory also needs to
integrate cyber security into its processes, including adding threat
model descriptions for control system design, and clean-slate designs
to improve resilience.

\subsection{Focus Areas}

\begin{enumerate}
	\item Control theory research to improve algorithms
	to be resilient against faults caused by cyber attacks
	\item Sensor correlation to survive owning of sensors
	\item Clean slate control theory design for resilience and
	fail-safe recovery (including graceful degradation and restart in the
	event of a cyber attack)
\end{enumerate}
\end{document}

### 2015-11-07 summary.tex
Inital thoughts on the tpcp projet

\documentclass[11pt]{article}

\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{url}
\usepackage[english]{babel}
\usepackage{vmargin}
\usepackage{times}
\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
\usepackage{cite}
\usepackage{titling}
\usepackage{color}

\widowpenalty=10000
\clubpenalty=10000

\begin{document}

\title{Executive Summary: Total Platform Cyber Protection}
\date{}

\maketitle

Traditional security defenses for the computing environment have
focused on security the \emph{border} of the network, as this is the
entry place for most attackers. However, even with sophisticated
border technologies in place, there are will a rash of security
breaches and data compromises. The goal of this work is to survey the
state of security for the entire computing platform, in order to
identify areas that require additional research.

In this work, we analyzed the security research done at the following
layers of the computing platform: hardware, firmware, bus, hypervisor,
operating system, application, network, and user layer.

At the hardware level, the threat is Trojan logic introduced during
design or manufacturing process. While much work has focused on
detecting Trojans using side-channel analysis, we believe that there
exists opportunity for developing the upper layers of the software to
mitigate the problem (assume untrusted hardware), detecting exploits
that utilize the Trojan, and to improve the detection process, both
before and after hardware is fabricated.

At the firmware level, the threat is that firmware has a high
privilege level and is opaque code that can easily hide a Trojan or a
vulnerability. We believe that more research is necessary on
automating the analysis of firmware, including reverse engineering,
transforming firmware, and increasing analysis to support the
concurrent and real-time requirements.

At the bus/interconnect layer, the threat is that any device attached
to the bus can communicate or impersonate any other device on the bus.
While current approaches focus on application-specific monitoring, we
believe that there must be more research on application-agnostic
monitoring, runtime bus filtering/policy enforcement, and device
authentication.

At the hypervisor layer, the threat is that vulnerabilities in the
hypervisor itself can allow a malicious or compromised guest operating
system to exploit the hypervisor and migrate to other guest operating
systems. While much research has been done on this layer, we believe
that there exists significant opportunity to automatically reduce the
attack surface of the hypervisor, increasing assurance of the
hypervisor, and verifying the hypervisor implementation.

At the operating system layer, the issue is that malicious application
can target the operating system to exploit vulnerabilities.

\end{document}

### 2016-01-11 paper.tex
Updated version

@@
-\documentclass[11pt]{article}
-
+\documentclass[11pt,letterpaper]{article}
+\usepackage[pass]{geometry}
 \usepackage{graphicx}
@@
 Traditional security defenses for the computing environment have
-focused on securing the \emph{boarder} of the network, as this is the
-entry place of most attacks. Specifically, technologies such as
-firewalls, intrusion detection systems, and automated threat forensics
-operate at the network gateway, attempting to effectively secure the
-boarder of the network.
+focused on securing the \emph{border} of the network, as this is the
+entry place for most attackers. However, even with sophisticated
+border technologies in place, there are constant attacks and data
+breaches against our networks. The goal of this work is to survey the
+state of security for the entire computing platform, in order to
+identify specific areas that are underserved by the research community
+and that require additional research and investment.
 
-However, even with such technologies in place, there is a rash of
-security breach and compromises. Target. OPM. Trump Industries.
+Modern computing infrastructure serves a wide array of needs, with
+diverse technologies and implementations: embedded systems, cloud
+computing, sensor networks, desktops, mobile devices, and industrial
+control systems. Rather than consider each of these computing
+platforms independently, in this work we abstracted the computing
+platform into several layers, each of which are applicable to every
+computing infrastructure. In particular, we analyzed the security
+research performed at the following layers of the computing platform:
+hardware, firmware, bus, hypervisor, operating system, application,
+and network layers.
 
@@
 machine, potentially attacking the underlying computing layers of the
-same machine or attacking other machines. Therefore, if we wish to
-increase the security of our computing systems, and reduce the number
-and scope of security breaches, it is essential that we encourage
-focus on novel ideas, algorithms, and techniques to secure every level
-of the computing stack.
+same machine or leveraging their place in the network to attack other
+machines. Therefore, if we wish to increase the security of our
+computing systems, and reduce the number and scope of security
+breaches, it is essential that we encourage focus on novel ideas,
+algorithms, and techniques to secure every level of the computing
+stack.
 
-In this paper, we will first discuss the computing stack, and, for
-every layer, we will identify focus areas that attempt to secure that
-layer. For each of those focus areas we will answer the following
-questions:
-
-\begin{itemize}
-  \item What's the technical challenge for the focus area?
-  \item Why is it hard?
-  \item What are the open problems?
-  \item How is the area addressed in the past (from a cyber perspective)?
-  \item Are there intellectual tools that need to be brought to bear on the problem?
-  \item How/when do we achieve results?
-  \item Near-term and the long-term?
-\end{itemize}
+In this report, we will first discuss the computing stack, and, for
+every layer, we will describe the layer, identify primary threats
+against the layer, and summarize the research on attempts to secure
+that layer, and finally we will determine areas that require further
+research investment to ensure the security of the layer.
 
@@
 
-What is the background of the computing layers? Here, having the
-layers in a graphic, explaining the interconnected layers would be
-good.
+Here, we consider the diverse array of computing platforms as a number
+of abstracted layers: hardware, firmware, bus, hypervisor, operating
+system, application, and network layers. In this way, we can
+categorize the research efforts that target each layer.
+
+The layers are arraigned from lowest to highest layer. The lower layer
+has more control over the computing system, while the upper layers
+have less control. For instance, the Hardware layer has total control
+over the physical hardware and memory of the machine, while an
+application running in the Application layer can only see a portion of
+the machine's memory (because of the virtual memory used by the
+Operating System).
+
+A typical compromise, in relation to the proposed layers, is the
+following: A vulnerability in an application is targeted, thus
+allowing the attacker full control of the application layer. From
+here, the attacker can scan the other applications on the network and
+attempt to exploit those applications, thus spreading horizontally
+throughout the network. Alternatively, as the application layer is the
+least privileged and least persistent of the layers, the attack can
+attempt to further the compromise on the lower layers of the machine.
+For instance, the attacker can exploit the hypervisor layer to
+compromise the host hypervisor. From this vantage point, the attacker
+can transparently extract information from all the other guest
+Operating Systems. More insidiously, the attacker can even exploit the
+firmware level, which allows the attacker to persist even after
+a complete reinstall of the operating system.
+
+In our abstract model, the lower levels have more control over the
+computing system, yet they have a smaller attacker surface. However,
+the lower levels can be attacked by the upper levels. Also, the upper
+levels trust the abstractions provided by the lower levels. For
+instance, the application-layer visualization of an industrial control
+system trusts that the data reported by the hardware layer it
+accurate. If an attacker is able to control and manipulate the
+hardware to falsely report the state of the system, the entire
+integrity of the system can be compromised.
+
+The interconnected nature of the layers means that the security of the
+entire layers must be a priority. However, at the same time, the
+security improvement of any layer will have an amplifying affect of
+the security of the whole system. For instance, improvement to the
+security of the application layer decreases the chance of an
+application compromise, which further decreases the risk of the attack
+propagating throughout the network and through the other layers.
 
 \section{Hardware}
+
+Hardware underpins all of modern computing. While most software is
+constructed and written in such as way as to abstract the
+peculiarities of the underlying hardware, the software could not
+execute without hardware. In our abstracted model of the computing
+stack, hardware is at the bottom, which means that it has complete
+control of the computing system.
+
+For hardware here, we consider all types of hardware in the computing
+system: CPUs, SoCs, or any other integrated circuit (IC) that performs
+computation of communication on behalf of the upper computing layers.
+This also include networking cards and other peripheral devices
+(Bluetooth or USB devices), which may have their own special purpose
+embedded hardware.
+
+A key difficult in analyzing the security concerns of hardware is
+reverse engineering the workings of the hardware. Traditional
+techniques to reverse engineer hardware require destroying the chip,
+and if a single hardware component is destroyed to reverse engineer,
+there is no guarantee that other hardware is identical to the reverse
+engineered chip.
+
+\subsection{Threats}
+
+As the lower-most level of the computing stack, all layers implicitly
+trust that the hardware is executing correctly. Therefore, if an
+adversary is able to maliciously control the hardware execution, they
+can compromise all layers of the computing stack. Thus, the security
+of the hardware is of paramount importance.
+
+The main threat considered by the literature on hardware is the threat
+of Trojan logic. This is malicious logic that is introduced into the
+hardware chip by a malicious adversary. Trojan introduction can be
+done at several points of the hardware lifecycle:
+
+\begin{itemize}
+
+\item Design. A malicious insider can incorporate a Trojan into the
+  design of the hardware.
+
+\item Third-party IP. Oftentimes, hardware designers will take
+  advantage of the functionality of third-party components, using them
+  as a black-box. Thus, a third-party can include malicious Trojan
+  logic.
+
+\item Compilation of Design. When the design of the hardware is
+  compiled, but before it is sent for fabrication, the design could be
+  changed to incorporate the Trojan logic. Note that this differs from
+  Design-time, because the Trojan logic is not included in the design.
+  Consider the case of malware on a hardware designer's machine, which
+  changes the design when it is compiled.
+
+\item Manufacture. A malicious manufacturer can introduce Trojan logic
+  into hardware after the design is received but before the hardware
+  is fabricated.
+
+\item Delivery. A malicious hardware component containing Trojan logic
+  can be swapped with a benign component after fabrication but before
+  delivery and installation.
+
+\end{itemize}
+
+The malicious Trojan logic behavior comes in three different forms:
+
+\begin{itemize}
+
+\item Trigger-based. The malicious behavior or logic is only expressed
+  when certain conditions are met. Consider a Trojan logic in a CPU,
+  which changes the executing code to Ring 0 (with the most
+  privileges) when a specific sequence of instructions are issued. The
+  trigger can be internal to the chip, such as in the CPU case, or
+  could be external, for instance a certain radio frequency will
+  trigger the malicious code.
+
+\item Constant. The malicious behavior or logic is continually
+  executed and subtly changes the output of the hardware in a way that
+  compromises security. Consider a Trojan logic of a Wi-Fi card, which
+  leaks the Wi-Fi passphrase by embedding it in the operating
+  parameters of the Wi-Fi radio.
+
+\end{itemize}
+
+These features of Trojan logic in hardware drive the research that
+attempts to detect them.
+
+\subsection{Research}
+
+Most research into hardware Trojans deal with using side-channel
+analysis to identify the malicious Trojan behavior~\cite{Lee2005,
+  Agrawal2007, Banga2008, Banga2008a, Jin2008, Du2010, Narasimhan2011,
+  Narasimhan2012, Cao2013, Davoodi2013, Dworak2013, Forte2013,
+  Liu2013, Liu2014, Bao2015}. This research uses power analysis, heat
+analysis, or other side-effects to determine that the hardware is
+executing abnormal logic. Unfortunately, this typically requires
+building a model based on a ``golden'' or known-good hardware, which
+is often not accessible. Furthermore, as chips become more complex,
+the Trojan logic become an increasingly smaller portion of the chip,
+which in turn reduces the side-channel behavior of the chip
+significantly.
+
+Other research has demonstrated the attacks that are possible by
+actually implementing Trojan behavior~\cite{King2008, Jin2010,
+  Sturton2011, Liu2013, Jin2014, Zhang2014}. This line of research
+shows that Trojan logic is feasible for an attack.
+
+Of particular practical interest is a description of \emph{actual
+  Trojan logic} discovered in a the ProASIC3 Flash FPGA military
+chip~\cite{Skorobogatov2012}. While in this case it was not proved
+that the logic introduced to the chip was intentionally malicious,
+nevertheless there existed functionality in the hardware that was
+explicitly denied by the hardware designer.
+
+Another avenue of research is to change the design of the chip itself,
+in order to improve the testability of the chip or to increase the
+effectiveness of the side-channel analysis~\cite{Chakraborty2008,
+  Li2008, Salmani2009, Hicks2010, Rosenfeld2011, Waksman2011,
+  Bhunia2013, Davoodi2013, Rajendran2013, Rolt2014}. This line of
+research attempts to rethink designing the hardware itself. However, a
+major drawback is that this can increase the cost of the hardware and
+it must be somehow mandated to be used in COTS components.
+
+JTAG is a main avenue of reverse engineering hardware, so some work
+has investigated the problem of securing JTAG access to
+hardware~\cite{Rosenfeld2010, Ren2015}. The main idea is to allow only
+authenticated users to have access to the JTAG functionality, however
+this increases the cost and complexity of implemented JTAG.
+
+Another approach is to change the design of the chip to support
+runtime monitoring~\cite{Waksman2010}. In this way, the hardware
+itself is is monitoring itself, rather than external analysis of
+side-channels. However, this has all the same drawbacks of changing
+the design of hardware, in addition to not being robust in the face of
+a powerful attacker who is able to alter the design.
+
+Recent approaches suggest to use machine-learning techniques to
+identify the Trojan logic~\cite{Haider2015}. Another approach attempts
+to use a formal verification technique in order to prove that no
+malicious logic was added to the hardware~\cite{Guo2015}.
+
+Finally, there are several surveys conducted on hardware security,
+which provide a good overview of the field~\cite{Wang2008,
+  Tehranipoor2010, Guin2014, Guin2014a, Guin2014b}.
 
@@
 
-Identification and signatures using features that are more robust to
-aging/use over time
+\subsection{Suggested Focus Areas}
+
+A key problem with hardware security is non-destructive IC
+examination, identification, and authentication. How to ensure that
+the chip that is designed is what is actually fabricated, with no
+additional logic.
+
+An additional neglected area is how to detect malicious Trojan logic
+in third-party components. This would require some type of static
+analysis or simulated dynamic analysis of third-party components,
+which could detect Trojan logic.
+
+Another approach that could be effective is how to properly test the
+functionality of hardware after it has been fabricated. This could
+include changing the design of the chip to allow better/more guided
+fuzz testing of the chip itself, in order to discover the Trojan
+logic.
+
+Finally, a prevention approach could be taken, to monitoring the upper
+layers of the computing stack for attempts to exploit the underlying
+hardware. This can also include changing or altering the firmware
+layer in such as way that it is not possible to exercise the Trojan
+logic, potentially using randomization techniques similar to ASLR.
+
+\section{Firmware}
+
+Firmware is the layer of software that sits and interacts directly
+with the hardware. Typically, the firmware is controlled from the
+upper layers by device drivers in the Operating System.
+
+Many different hardware contain firmware, for instance the boot BIOS
+on a motherboard is a form of firmware, embedded systems often contain
+specific firmware, hardware peripheral such as network cards contain
+firmware as well. We consider all such areas as firmware.
+
+The key traits of firmware is that, although it is binary code, it is
+operating/interacting directly with the hardware, so there are no
+OS-style abstraction layers such as virtual memory, well-defined
+system calls, or hardware abstraction. Therefore, each firmware is
+binary code with is specifically tailed to a specific hardware
+platform.
+
+Due to the hardware-specific nature of firmware, \emph{reverse
+  engineering} is critical to understanding the firmware. In addition,
+reverse engineering is particularly difficult because source code is
+frequently unavailable, and, vendors desire to protect their
+intellectual property, therefore, they actively obfuscate their
+firmware. All this means that, extracting semantically useful
+information from firmware is a key first step to performing any
+security analysis of the firmware.
+
+\subsection{Threats}
+
+As the lowest-level of software code in the computing stack, threats
+to firmware can compromise the security of the entire machine.
+Firmware security is concerned with two main threats: (1)
+vulnerabilities in the firmware code itself which allow an attacker to
+execute arbitrary code with the capabilities of the firmware and (2)
+firmware with intentional Trojan (backdoor) capabilities.
+
+Vulnerabilities in the firmware code allow an attacker to execute
+arbitrary code with the permissions of the firmware. The class of
+vulnerabilities are not unique to firmware, as they are the same class
+of vulnerabilities that infect traditional binary software (buffer
+overflows, logic flaws, etc.). However, firmware typically has full
+control over the entire computing system, therefore vulnerabilities in
+firmware are much more severe than those in the application layer. In
+addition, if an attacker is able to persist their exploit, because the
+compromise occurs at such a low layer of the computing system, the
+attacker can persist \emph{even after reinstalling the operating
+  system or application}.
+
+The attack surface for firmware varies depending on where it is used.
+In the most general case, the attacks can come from the upper layers,
+specifically the Operating System or Hypervisor layers, as these are
+the layers of the computing stack that can directly communicate to the
+firmware. However, other firmware can be attacked either directly
+through the Hardware layer itself, or through the data being
+transmitted through the firmware from the hardware. For instance, a
+network card firmware could be attacked through network traffic that
+the firmware processes.
+
+Traditional vulnerability analysis is not effective on firmware, due
+to the specificity of the firmware for the specific hardware and the
+lack of a well-defined operating system abstraction for the firmware
+code. In essence, each firmware is a custom binary with little in the
+way of assumptions that vulnerability analysis can take advantage of.
+
+Trojan capabilities in the firmware itself can allow a system to be
+remotely controlled or accessed. This is similar to a vulnerability,
+in that an attacker is able to control the firmware, however the key
+difference lies in the fact that Trojan behavior will be intentionally
+obfuscated by an attacker, whereas a vulnerability is accidentally
+introduced by a benign developer. Therefore, Trojan behavior is more
+more difficult to detect, because there is an active adversary who is
+intentionally obfuscating the behavior.
+
+Vulnerabilities in the firmware and Trojan capabilities are related
+because both require \emph{understanding} the behavior and semantics
+of the firmware code. This is why reverse engineering is such a
+critical component of securing the threats against the firmware layer:
+in order to detect either vulnerabilities or Trojan functionality, an
+automated tool must first understand the semantics of the particular
+firmware/hardware combination.
+
+\subsection{Research}
+
+Researchers have attempted to identify real-world attacks on firmware
+code~\cite{Tsow2006, Cui2010, Duflot2011, Basnight2013, Cui2013,
+  Zaddach2013, Zaddach2013a, Maskiewicz2014, Stuttgen2015}. These
+attacks are particularly effective, and researchers have shown that
+these are able to persist through a reboot and cannot be detected by
+the upper layers.
+
+Other approaches have looked at how to prevent the tampering of
+firmware~\cite{Adelstein2002, Zhou2007, Zhou2009}. This protects from
+the threat of tampering or altering the firmware, however it does not
+address vulnerable firmware.
+
+Another approach has been to include run-time monitoring into the
+firmware, so that compromises can be detected~\cite{Cui2010a,
+  Cui2011}. This approach is further complicated by the lack of
+abstractions in the firmware environment.
+
+Recent work has looked at the problem of updating the design of the
+firmware so that it has security features, such as memory handling and
+avoiding information leakage~\cite{Koeberl2014, Kauer2007}. However,
+forcing firmware developers to use such features is an open issue.
+
+Very few work has looked at how to reverse engineer and understand the
+firmware~\cite{Zaddach2014}. This problem is of particular interest,
+as understanding the semantics of the firmware and how it interacts
+with hardware is a key challenge.
+
+Recent work has looked at how to detect vulnerabilities in
+firmware~\cite{Davidson2013, Costin2014, Shoshitaishvili2015}. One of
+these has an excellent website\footnote{\url{http://firmware.re}}
+that contains the results of their reverse engineering and
+vulnerability analysis~\cite{Costin2014}.
+
+\subsection{Suggested Focus Areas}
+
+Of all the layers, firmware has received the least attention from the
+research community. Therefore, we believe that focus on this research
+area will yield significant dividends.
+
+The major issue with firmware vulnerabilities is how to decouple
+firmware from the hardware, potentially using emulation of the
+hardware, in order to aid the analysis of the firmware.
+
+In a similar vein, there is need for automated reverse engineering tools,
+including analysis of memory layouts, code modules, even answering the
+simple question of what in a firmware binary is code and what is data.
+
+Finally, even if we find vulnerabilities or Trojan logic in a
+firmware, there is still the unanswered questions of how to patch or
+update the functionality of the firmware while still guaranteeing the
+behavior of the firmware. This is particularly complicated due to the
+real-time requirements of the firmware and the potentially concurrent
+executions.
+
+\section{Bus and Interconnect}
+
+The Bus and Interconnect is an abstract layer of the computing system
+which allows the firmware and hardware to communicate with each other.
+Different computing architectures have different technical
+implementations of the bus and interconnect. For instance, the
+communication network in the automotive system, which has received
+significant attention from the research community, uses a CAN bus to
+communicate. Modern motherboards us either a Northbound or Southbound
+bus to communicate between the CPU and the other peripheral devices,
+such as USB.
+
+\subsection{Threats}
+
+Depending on the technical architecture of the bus and interconnect
+layer, the bus can be used either purely for communication or for
+other functionality as well. For instance, the CAN bus is only used to
+allow different hardware components to communicate. However, on modern
+motherboards, techniques such as Direct Memory Access (DMA) allow a
+peripheral device that is connected to the bus to directly access the
+system's memory, without going through the CPU.
+
+In the case of the bus and interconnect being used as a communication
+medium, then the classic threats are the same security concerns as
+other mediums of communication: communication integrity and
+confidentiality. A malicious device on the bus can impersonate another
+device and send messages on the bus as the impersonated device. On an
+automotive system, researchers have demonstrated that by compromising
+a single device attached to the system's CAN bus allows an attacker to
+have complete control over all functions of the automobile, by
+impersonating messages on the CAN bus. In addition to impersonating
+messages, an attacker on the bus and interconnect layer can listen to
+messages on the bus, potentially eavesdropping on sensitive messages.
+
+If the bus and interconnect layer offers advanced functionality, such
+as the DMA example, then devices that an attacker is able to connect
+to the bus could take advantage of this functionality. In the case of
+DMA, this means that any device on the bus has full access to the
+physical memory of the system, therefore the attacker has complete
+control over the upper layers of the computing system (hypervisor,
+operating system, and application).
+
+The difficulty of addressing the threats of the bus and interconnect
+layer is the diversity and specificity of the specific bus and
+interconnect technology. Often, these were created without
+consideration for security needs, therefore, as a communication
+mechanism they lack authentication from the design. In addition, as
+the bus and interconnect layer cause different hardware to
+communicate, all hardware must agree on the protocol to talk on the
+bus. Therefore, adding security to the bus and interconnect protocol
+is difficult in legacy systems.
+
+\subsection{Research}
+
+A major area of research in the Bus and Interconnect layer is work
+that shows the danger of a bus and interconnect that allows malicious
+devices to either impersonate devices on the bus or leverage bus
+capabilities, particularly in the automotive area~\cite{Bonkoski2013,
+  Checkoway2011, Hoppe2008, Koscher2010, Rouf2010, Sang2010}.
+
+The other major area of research is to apply encryption to devices on
+the bus and interconnect~\cite{Groza2012, Herrewege2011, Jiang2012,
+  Schweppe2011, Stewin2014, Szilagyi2010}. In this way, malicious
+devices on the bus cannot impersonate other devices. However, a key
+problem with this research direction is how to distribute the keys of
+the devices, which is typically done by distributing them in hardware.
+Unfortunately, this approach is brittle, as it disallows new devices
+from accessing the bus and interconnect.
+
+Other research has investigated how to ensure the integrity of devices
+attached to the bus and interconnect using attestation
+techniques~\cite{Li2010, Li2011}. These techniques attempt to ensure
+that only authorized and unmodified devices are attached to the bus
+and interconnect. However, this research has the problem of how to
+allow third-party devices to access the bus.
+
+A key problem is how to perform secure updates to devices attached to
+the bus and the interconnect~\cite{Larson2008}. This is a challenge
+that also has similarities with the firmware layer.
+
+In order to discover malicious communication, researchers have
+utilized data flow tracking techniques to discover how the data flows
+between devices connect to the bus and interconnect
+layers~\cite{Schweppe2012}. Using these techniques, the researchers
+are able to identify malicious flows or potential vulnerabilities by
+analyzing the way that data flows throughout the system.
+
+Other researchers have explored modeling either the system itself
+or the behavior of the devices on the bus~\cite{Stewin2013a,
+  Drolia2011}. If the device behavior or the system behavior does not
+match the learned model, then an alert is triggered. In this way,
+these techniques are similar to anomaly detection techniques, with
+similar benefits and drawbacks.
+
+In order to address the issue of devices on the bus able to access
+capabilities unauthorized, researchers have tackled the similar
+problem of reducing the Trusted Computing Base, essentially reducing
+the functionality of the bus and interconnect
+layer~\cite{Vasudevan2012, Zhang2013, Zhou2009}. In this way, if the
+functionality is not there on the bus, then a malicious device cannot
+take advantage of it.
+
+Another approach is to leverage hardware features of the underlying
+bus and interconnect system to reduce the impact of an
+attack~\cite{Wolf2012, Wojtczuk2011}. However, as these hardware
+functionality must be implemented, it's impact in legacy systems is
+limited.
+
+Finally, other research in this area has performed surveys of the
+research on this area~\cite{Kleberger2011, Studnia2013, Thom2008,
+  Wolf2004, Wolf2007, Wright2011, Zhao2002}.
+
+\subsection{Suggested Focus Areas}
+
+Further research in this area should focus on enabling better
+monitoring of the runtime of devices attached to the bus or
+interconnect. There is also the need to support logging of the bus or
+interconnect activity. This will enable to detection of a malicious
+device on the bus or interconnect.
+
+Another promising area of research is into the automated filtering of
+the bus communication, potentially allowing an administrator to define
+a policy for a specific bus or interconnect system. This would be
+particularly effective if it can be implemented on legacy bus and
+interconnect systems, with no modification to the devices.
+
+A different approach would be to secure the bus by only allowing
+authorized devices to attach to the bus. This could be a white-listing
+based approach or other policy-driven approach. However, the key
+question is how to identify a device, particularly a third-party
+device.
+
+While there has been some research on encrypting the communication on
+the bus, most has focused on pre-distributing the keys to devices on
+the bus. This approach is brittle, as new devices cannot be added to
+the bus easily. Therefore, further approaches should be investigated,
+such as using attribute-based encryption to be able to encrypt and
+distribute messages on the bus without prior sharing of keys.
+
+\section{Hypervisor}
+
+The purpose of the hypervisor layer is to provide virtualization
+techniques for the upper layers of the computing stack. In essence,
+the Hypervisor layer provides a virtualization or emulation of the
+lower layers of the computing stack, so that multiple Operating
+Systems, which each assume that the are the sole users of the
+underlying hardware, can run concurrently on the same hardware.
+
+For discussing the Hypervisor layer, we use the term \emph{host} to
+define the hypervisor itself (which, in some cases can be an Operating
+System) and the \emph{guest} is the Operating System that is running
+in the hypervisor. A host can run multiple guest Operating Systems.
+
+Hypervisors have a number of benefits, from allowing greater sharing
+and usage of hardware resources, to improving the isolation of the
+Operating Systems. The security benefits of using Hypervisors comes
+from each guest Operating System being isolated from the other. In
+theory, this means that a compromise of a guest Operating System will
+be isolated to only that Operating System, and the attack should not
+spread to other guest Operating Systems.
+
+\subsection{Threats}
+
+The central threat to the Hypervisor layer is from attackers in a
+guest Operating System exploiting a vulnerability in the hypervisor
+itself to gain access to the actual hardware or to propagate their
+attack to the other guest Operating Systems that are running in the
+hypervisor. As the Hypervisor layer is essentially emulating the
+hardware for the guest operating systems, it has a large attack
+surface. Furthermore, modern hypervisors, in an attempt to increase
+the performance of the hypervisor, frequently allow guest Operating
+Systems direct access to the hardware. If the hardware does not
+provide for full isolation (i.e., it is not hypervisor-aware), then an
+attacker can leverage this hardware to tamper or extract sensitive
+information from other guest Operating Systems.
+
+In modern hypervisors, the hypervisors frequently provide isolation
+from a well-behaved guest Operating System, as the guest Operating
+System is assumed to be benign. This assumption can be unfounded, as
+the guest Operating System can be hostile and can attempt to exploit
+or crash the hypervisor.
+
+Another threat actually stems from the Hypervisor itself being
+untrusted. In the cloud computing environment, the guest Operating
+System does not trust the underlying Hypervisor, as it is controlled
+by a third-party. In this case, the Hypervisor can spy on the guest
+or alter the memory of the guest.
+
+\subsection{Research}
+
+One area of research is to leverage the vantage point of the
+Hypervisor to monitor the guest for security
+violations~\cite{Karger1991, Barham2003}, intrusion
+detections~\cite{Dunlap2002, Garfinkel2003a, Kourai2005, Chen2008,
+  Garfinkel2003b}, checking the integrity of the guest Operating
+system~\cite{Seshadri2006, McCune2010}, and identifying rootkits present
+in the guest Operating System~\cite{Jones2008, Grace2010}. This work
+assumes that the guest can be compromised and that the Hypervisor is
+trusted, which, due to vulnerabilities in the Hypervisor, is not
+always true.
+
+A significant problem in monitoring or otherwise inspecting a guest
+Operating System from the hypervisor layer is the \emph{semantic gap}
+problem, where the Hypervisor has no outside knowledge of the
+guest-specific Operating System constructs. Therefore, research has
+investigated how to overcome this semantic gap~\cite{Jones2006,
+  Christodorescu2009, Srinivasan2011}.
+
+Further work in this line has approached the problem of how to monitor
+the hypervisor itself by leveraging hardware features~\cite{Azab2010}.
+
+In order to build a secure Hypervisor, the lower layers of the
+computing system must also be trusted. Therefore, research into
+establishing a Trusted Computing Base, which validates the integrity
+of the lower levels for the guest Operating System~\cite{Arbaugh1997,
+  Sailer2004, Singaravelu2006} is important for ensuring the security
+of the Hypervisor.
+
+One problem of using the Hypervisor as a security primitive is that it
+is part of the Trusted Computing Base for the upper layers of the
+computing platform. Therefore, attempts have been made to reduce the
+trusted base of the Hypervisor itself~\cite{Murray2008, Shinagawa2009,
+  Colp2011, Wang2012}.
+
+Other research has looked at creating a secure
+Hypervisor~\cite{Robin2000, Belay2012, Szefer2011, Vasudevan2013} with
+the possibility of adding functionality to the Hypervisor to support
+closed guest Operating Systems~\cite{Garfinkel2003}, offering
+applications running in the guest Operating System to hide information
+from the guest Operating System~\cite{Chen2008, Yang2008, Jin2011,
+  Szefer2012, Azab2011, Zhang2011, Jin2011a, Szefer2011a}, secure
+information flow between the guest Operating System drivers and I/O
+devices~\cite{Cheng2013}, or designing the hypervisor so that it can
+prove that it did not tamper with the guest Operating
+System~\cite{Gu2011}.
+
+Another approach to improve the security of the Hypervisor is to
+remove the duplicated components from the guest Operating System
+layer~\cite{Kivity2014, Madhavapeddy2015}, which effectively reduces
+the attack surface of the Operating System layer.
+
+Another interesting approach is to run traditional network middleware
+boxes, such as firewalls and Intrusion Detection Systems, as a guest
+Operating System of a Hypervisor~\cite{Martins2014}.
+
+Finally, there have been attempts to introduce traditional
+application-level security features to Hypervisors, in order to reduce
+or prevent vulnerabilities, such as control-flow
+integrity~\cite{Wang2010} or manual vulnerability
+analysis~\cite{Perez-Botero2013}.
 
@@
 
-\begin{enumerate}
-	\item Non-destructive IC examination, identification, and
-	authentication
-	\item Continuous or near-continuous physical monitoring for
-	establishing root of trust
-	\item Inspection-friendly design for hardware transparency
-\end{enumerate}
+As the hypervisor is typically a monolithic trusted entity, research
+should be performed to reduce the trust in the hypervisor. One avenue
+could be to have fine grained access control within the hypervisor, so
+that different policies can be set to the guests hosted by the
+hypervisor.
 
-\section{Firmware}
+In a similar area, we believe that further research into reducing the
+attack surface of the hypervisor is necessary. Specifically, the
+hypervisor can be customized based on the applications and workload
+necessary.
 
-Firmware must deal with the fact that no OS-style abstraction layer
-is available, so hardware interaction is important.  Reverse
-engineering is an interesting at the firmware level because most
-vendors want to make it difficult to protect IP; it is also more
-tedious and would the RE process would benefit more from better tools.
+Hypervisors, despite their obvious benefits, have not been applied to
+real-time systems. Further research must be done to allow reliable and
+robust hypervisors in a constrained real-time computing environment.
+
+Testing hypervisors is currently a difficult and essentially manual
+process. There is currently no equivalent to fuzz testing a hypervisor
+implementation, possibly because there are no models for attacks
+against a hypervisor. Further research in increasing the testability
+and automated analysis of hypervisor reliability would improve security.
+
+\section{Operating Systems}
+
+The Operating System layer is responsible for ensuring that multiple
+applications are able to concurrently run, without interfering with
+each other. In addition, the Operating System layer is also
+responsible for providing a unified abstraction of the underlying
+hardware layer for the application layer. Therefore, in terms of
+security, the Operating System layer has more privileges than the
+application layer, as the Operating System controls the execution and
+scheduling of processes, the access control policies that allow one
+application to hide information/data from the other processes.
+
+\subsection{Threats}
+
+The main threat to the Operating System layer is an attacker
+compromising the security of the Operating System. There are two main
+avenues to how this compromise can occur: (1) an attacker exploits a
+vulnerability in the Operating System from the Application layer or
+(2) an attacker compromises a lower layer (for instance, modifying the
+Operating System code directly to the hardware's Hard Drive).
+
+Vulnerabilities in the Operating System allow an attacker to have
+complete control over the other application running in the operating
+system. In addition, the attacker can use their vantage point at the
+operating system to attempt to escalate their attack to the hypervisor
+or lower layers, such as the firmware. Vulnerabilities in the
+Operating System result from programming errors that allow an attacker
+to gain control of the Operating System code or data. These
+vulnerabilities are traditional memory corruption vulnerabilities or
+even logic flaws.
+
+A main concern in Operating System security is the presence of
+rootkits, which are malicious modifications to the Operating System
+which an attacker uses to achieve persistence on the system. In
+addition, attackers use rootkits to hide from other applications on
+the system. For instance, the rootkit will modify the operating system
+code such that the attacker's processes and files are hidden from the
+administrator or defensive software installed on the Operating System,
+such as an anti-virus solution.
+
+Direct access to the hardware can allow an attacker to compromise the
+Operating System. It is in this way that an attacker can propagate
+their compromise from the lower layers of the computing platform to
+the upper layers. The attacker has similar capabilities as when
+exploiting a vulnerability in the Operating system.
+
+\subsection{Research}
+
+Research into Operating System security has a long and rich history,
+which has looked at design flaws in Operating
+Systems~\cite{Attanasio1973}. Further research looked at reducing the
+permissions of the \emph{root} users, essentially compartmentalizing
+the privilege of the Operating System~\cite{Kamp2000, Kurmus2014a}.
+
+Rootkits attempt to persist the attacker's compromise of the operating
+system and allow them to hide their malicious behavior from other
+applications on the system. Therefore, rootkit detection is an
+important research topic in Operating System
+security~\cite{Petroni2004, Wang2005, Petroni2007, Payne2008,
+  Sharif2009, Yin2010, Bianchi2012}. Another approach is to prevent
+the rootkit from altering the Operating System
+kernel~\cite{Riley2008a, Wang2009, Grace2010, Gadaleta2011}.
+
+Vulnerability analysis techniques have been used to discover
+vulnerabilities in the Operating System kernel before a hacker is able
+to use and exploit those vulnerabilities. Static analysis techniques
+have been used~\cite{Chou2001, Chen2011}.
+
+Other vulnerability defenses include defenses for specific
+vulnerability classes, such as buffer overflows~\cite{Dalton2008}.
+Randomization techniques attempt to alter the environment of the
+operating system in such as way that the system continues to function
+but the attacker's exploit fails~\cite{Lin2009, Giuffrida2012}.
+Control-flow integrity techniques have been applied to Operating
+System code as well~\cite{Li2011a}.
+
+Other research into Operating System security has looked at how to
+protect applications from a malicious or compromised Operating
+System~\cite{Ta-min2006, Onarlioglu2013, Li2014}.
+
+As driver code executes in kernel space, and are a frequent source of
+vulnerabilities~\cite{Kadav2012}, researchers have looked into how to
+make drivers more robust~\cite{Ganapathy2008, Hruby2012} or isolated
+from the rest of the kernel~\cite{Boyd-Wickizer2010}.
+
+Attack papers show that it is possible to compromise the kernel
+without injecting any additional code, using Return Oriented
+Programming techniques~\cite{Hund2009}.
+
+A completely different approach is to formally prove the correctness
+of the Operating System kernel~\cite{Klein2009}.
+
+Updating Operating Systems with security updates is frequently
+difficult, as the Operating System must be restarted. However,
+researchers have explored how to apply an update to the Operating
+System without rebooting the system~\cite{Arnold2009}.
+
+Some work has been done on removing the operating system
+entirely~\cite{Madhavapeddy2010} or trimming/tailoring parts of the
+Operating System~\cite{Kurmus2011, Madhavapeddy2013, Howell2013,
+  Kurmus2013, Kurmus2014}. This has the security benefit of reducing
+the attack surface of the Operating System.
+
+\subsection{Suggested Focus Areas}
+
+We believe that, while significant research has been done on the
+Operating System layer, there still remains work to be done on this
+area. In particular, as the Operating System is frequently in contact
+with malicious applications, it must be hardened from attack.
+
+Therefore, we believe that further research into the line of trimming
+or tailoring the operating system is important, such that the attack
+surface is significantly reduced.
+
+We also believe that non-control data-flow attacks will be the new
+wave of attacks against Operating Systems. Therefore, we need new
+analysis techniques, both static and dynamic, to discover these
+vulnerabilities that do not require altering the control flow of the
+Operating System.
+
+A critical component of the Operating System is to handle the
+execution of concurrent applications. Therefore, attacks that take
+advantage of the concurrency in the Operating System are unexplored.
+Research must be done to create new models for these attacks.
+
+\section{Application Layer}
+
+The application layer is the layer of the computing stack that
+performs that actual useful computation or offers the computing
+service. It is essentially at the top of the computing stack, as the
+lower layers of the computing stack provide abstractions so that
+applications are easier to write and the upper layer (networking) is
+used by the application layer to communicate to other applications.
+
+At the application layer, we consider all the following types of
+``applications'' under the application layer umbrella: GUI
+applications, command-line applications, server batch process
+applications, web applications, or mobile applications. Also, all
+programming languages used to develop the applications are also in
+scope, from compiled languages such as C or C++ to scripting languages
+such as Python and Ruby. In short, all types of computing application
+are included in the application layer.
+
+Due to abstractions provided by the lower layers, applications
+typically have less privilege than the lower layers of the computing
+stack. This is due to the abstractions provided to the application
+layer, for instance an application cannot access the underlying hard
+drive directly, but must use the Operating System's provided
+interface.
+
+Applications developed are typically very complex to understand
+completely. This is due in part to the reliance on the abstractions
+provided by the lower computing layers (as these abstractions must be
+fully understood to understand how the application uses them). Another
+aspect is that applications frequently use shared libraries or other
+middleware, which as also considered to be part of the application
+layer. In this way, multiple different applications can all use the
+same underlying library. As the application's behavior depends on the
+behavior of \emph{all} used libraries and abstractions, this means
+that understanding even a simple application is dependent on the
+complexity of the used abstractions.
+
+\subsection{Threats}
+
+Applications, by definition must be accessible by users in
+order to perform their useful computation. Due to this accessibility,
+the application layer, of all the other computing layers, is the entry
+point for attacks or exploits.
+
+At a high level the threats to the application layers is an attacker
+able to influence the system to perform computation of their choosing.
+There are two main vectors for an attacker to influence the system:
+(1) software vulnerabilities in an application that can allow an
+attacker to exploit the vulnerability to either perform arbitrary
+computation or otherwise influence the application and (2) a malicious
+application (typically described as a Trojan application) that the
+attacker tricks the user into installing and running on the system,
+usually by using social engineering techniques.
+
+Vulnerabilities are flaws or bugs in the application (or any software
+for that matter) that allows an attacker to compromise the security of
+the application. The key difference from normal flaws or bugs in
+software is that vulnerabilities allow the program to deviate from the
+behavior. These vulnerabilities range the gamut of possible causes
+such as programming errors or configuration mistakes. They can be in
+the actual, custom application software or in any of the libraries
+that the application uses. Some vulnerabilities are specific to a
+given programming language, while others are more broad and can any
+affect application. Examples of common vulnerabilities are: buffer
+overflow, heap overflow, format string, off-by-one, integer overflow,
+time-of-check-time-of-use, cross-site scripting, SQL injection, and
+logic flaws.
+
+Trojan applications attempt to control or manipulate the system by
+tricking a user into executing them. They frequently do this by using
+social engineering techniques to entice the user to execute them. The
+application may appear innocuous, such as a flashlight application on
+a mobile device, but may in fact include malicious functionality that
+the user does not want.
+
+Using either of these two techniques, the attackers goal is to move
+from an external position to having influence on the executing system.
+The best scenario for the attacker is to completely control the
+execution of the application. From here, the attacker can now
+arbitrarily access any resource that the application can access. For
+instance, this could mean reading sensitive data, changing sensitive
+data, deleting sensitive data, or using the vantage point of the
+application to further the compromise to the lower levels of the
+computing platform or even to other applications on the network.
+
+Applications have a large and complex attack surface, which increases
+the chances of a developer accidentally introducing a vulnerability.
+The attack surface of the application is increased by the transitive
+closure of the libraries that the application uses (and that those
+libraries use). A vulnerability in a library is a vulnerability in the
+application that uses that library. Therefore, the complexity and
+increased attack surface of applications are a threat, as they allow
+increases in vulnerabilities.
+
+\subsection{Research}
+
+Security of the application lays has received significant attention
+from the research community. Because of this, we have divided the
+research into attack surface reduction, prevent control-flow
+hijacking, logic and implementation flaws, mitigation and recovery,
+modeling and monitoring execution, and protection against malicious
+code.
+
+\subsubsection{Attack surface reduction}
+
+The idea behind this subarea of application layer security research is
+to reduce the attack surface of the application. This line of research
+views vulnerabilities as a probability, and by reducing the attack
+surface, the odds of an attacking being able to exploit a
+vulnerability successfully is reduced as well.
+
+Attempts have been made to measure an application's attack
+surface~\cite{Howard2003, Manadhata2007, Manadhata2008, Ruprecht2014}.
+This has the benefit of attempting to quantify the attack surface of
+an application, which in itself is a difficult problem.
+
+Several research works attempt to reduce the attack surface. Some
+approaches try to statically analyzing for deadcode, then restrict the
+execution of the deadcode during runtime~\cite{Kurmus2011}. Other ways
+attempt to leverage compile-time configurability to reduce the attack
+surface~\cite{Tartler2012, Kurmus2013, Ruprecht2014}. A different
+approach is to reduce the attack surface by partitioning the
+application into different privilege levels~\cite{Barth2010, Xu2012}.
+Other approaches attempt to restrict the communication between
+applications, in order to reduce opportunities for the attacker to
+input data to the system~\cite{Kantola2012}.
+
+Finally, other approaches attempt to quantify the attack surface and
+measure the actual reduction~\cite{Kurmus2014}.
+
+\subsubsection{Prevent control-flow hijacking}
+
+Modern attackers, exploiting binary applications, such as browsers,
+media players, or PDF readers, attempt to hijack the control-flow of
+the application. If an attacker successfully hijacks the control-flow
+of the application, then she has complete control over the execution
+of the application. Many vulnerability classes enable control-flow
+hijacking, specifically memory corruption vulnerabilities such as
+buffer overflows, heap overflows, and format strings~\cite{Szekeres2013}.
+
+The goal of this line of research is to prevent the attacker from
+successfully hijacking the control-flow of the application. In this
+way, this line of research can be seen as a mitigation technique.
+
+Significant progress in this area started with the idea of
+control-flow integrity, essentially ensuring dynamically at run-time
+that the control-flow of the application does not
+deviate~\cite{Abadi2005, Abadi2009}. The first implementations were
+deemed infeasible, so later research attempted to improve the
+efficiency of control-flow integrity~\cite{Zhang2013, Tice2014,
+  Bletsch2011, Zhang2013, Zhang2013a}
+
+Researchers have advanced the state of attacks against control-flow
+hijacking. The most advanced of these is return-oriented programming
+(ROP), in which an attacker reuses bits of code from the original
+application, so that the attacker can hijack the control-flow of the
+application without introducing new code~\cite{Roemer2012,
+  Bletsch2010, Tran2011}. Researchers have also improved these ROP
+techniques~\cite{Checkoway2010, Buchanan2008, Kayaalp2012, Chen2012,
+  Goktas2014, Schwartz2011} and applied them to different
+architecture~\cite{Davi2012}.
+
+Other research has looked at defending ROP attacks~\cite{Pappas2012,
+  Davi2009, Davi2011, Onarlioglu2010, Gupta2013, Pappas2012, Lu2011,
+  Lu2011a, Homescu2012, Gokta2014, Cheng2014, Pappas2013, Snow2013,
+  Hiser2012, Wartell2012} or detecting ROP
+attacks\cite{Polychronakis2011}. Metastudies looked at the
+effectiveness of these defenses~\cite{Skowyra2013, Schuster2014}.
+
+Finally, other research has shown that even with perfect control-flow
+integrity, due to the impreciseness of statically creating the
+control-flow graph, an attacker can still hijack the control
+flow~\cite{Carlini2015, Schuster2015, Carlini2014, Davi2014}.
+
+\subsubsection{Logic and implementation flaws}
+
+Not all security vulnerabilities are the result of memory corruption
+or other programming language-based flaws. Another kind of
+vulnerability class are called logic (or implementation) flaws. These
+flaws occur when the logic of the application is faulty. Consider an
+ecommerce application that allows a coupon to be reapplied to an order
+multiple times, driving the price to zero. Or if a user is able to
+enter a negative quantity for the purchase of an item, and the
+application uses that negative quantity to refund the purchase price!
+
+Logic vulnerabilities are particularly difficult for automated
+vulnerability analysis because they are deviations from the
+system's specification and the system's implementation. Very often, a
+system's specification exists only in the minds of the developers.
+Therefore, logic flaws are specific to every application, and are not
+simply the result of faulty data flows as most other vulnerability
+classes are.
+
+Research in logic flaws has typically focused on detecting a certain
+type of logic flaw, rather than detecting logic flaws generally. Early
+work looked at detecting state violations in web
+applications~\cite{Cova2007}.
+
+Other classes such as preventing access control
+vulnerabilities~\cite{dalton09:nemesis} or statically detecting
+them~\cite{Sun2011}. Privilege escalation
+vulnerabilities~\cite{Monshizadeh2014}, data disclosure
+vulnerabilities~\cite{Muthukumaran2015}
+
+Parameter tampering vulnerabilities are another type of logic
+flaw~\cite{Bisht2010, Bisht2011}
+
+Execution After Redirect are another type of logic flaw
+vulnerabilities in web applications~\cite{Doupe2011, Payet2013}
+
+Single sign-on vulnerabilities~\cite{Zhou2014}.
+
+Work on detecting logic vulnerabilities directly started with
+attempting to infer security specifications of an
+application~\cite{tan08:autoises, Livshits2009a}. Other approaches dynamically
+execute the application to infer likely invariants, then static
+analysis attempts to find paths that violation those
+invariants~\cite{felmetsger10:logic}. Further work tried to find missing
+security (authorization) checks automatically, without prior knowledge
+of the checks~\cite{Son2011, Son2013}.
+
+Recent work has looked at detecting logic flaw vulnerabilities in web
+applications in a black-box manner---that is, no knowledge of the
+application's source code~\cite{Pellegrino2014}.
+
+\subsubsection{Mitigation and recovery}
+
+\subsubsection{Modeling and monitoring execution}
+
+\subsubsection{Protection against malicious code}
 
@@
 
-\begin{enumerate}
-	\item Decouple firmware from the hardware (emulation) to aid analysis
-	\item Better tools for automated reverse engineering, including
-	analysis of layouts, modules, determining code vs. data
-	\item How to patch firmware with analysis and guarantees about
-	system behavior
-	\begin{enumerate}
-		\item did you fix the vulnerabilities?
-		\item did you introduce new vulnerabilities?
-		\item did your changes disrupt the timing of the original firmware?
-	\end{enumerate}
-\end{enumerate}
+Much of the prior research focuses on just looking at memory
+corruption bugs, but we need to look at how the same sort of advanced
+protections can be applied to other classes of bugs, i.e., what is
+next? This includes: input validation, logic problems, leaking
+secrets, performance degradation, denial-of-service, command (e.g.,
+SQL) injection, and concurrency attacks.
 
-\section{Bus and Interconnect}
-
-\subsection{Focus Areas}
-
-Critical tech challenges:
-\begin{enumerate}
-	\item Lightweight device authentication that is bus-appropriate
-	\item Operational mode setting to determine allowed capabilities
-	for each device
-\end{enumerate}
-
-\section{Hypervisor}
-
-\subsection{Focus Areas}
-
-Critical tech challenges:
-\begin{enumerate}
-	\item Protect private information within a VM/guest from spying
-	and interception by the host, bootstrap image with private key
-	\item Attest memory, e.g., did my cloud provider put something in
-	to read memory?
-	\item Achieve predictable real-time scheduling in a lightweight
-	and secure hypervisor
-\end{enumerate}
-
-\section{Operating Systems}
-
-Discussion centered on reducing the attack surface by stripping out
-unnecessary and unneeded functionalities.
-
-\subsection{Focus Areas}
+It is clear that an large amount of research effort has looked at
+securing the application layer. However, even so, there are areas that
+we believe are underserved by the research community and provide a
+significant opportunity to solve.
 
 \begin{enumerate}
-	\item Automated reduction and pruning of features to a minimal set
-	of required functionality
-	\item Non-control data flow attacks
-	\item Concurrency-based attacks
+	\item The next generation of vulnerabilities (e.g., logic flaws)
+	\item Need good tools for analysis of race conditions / concurrency attacks
 \end{enumerate}
-
-\section{Middleware / Runtime}
 
@@
 
-\section{Application Layer}
+%% \section{Network Layer}
 
-Much current work focuses on just looking at memory corruption bugs, but
-we need to look at how the same sort of advanced protections can be
-applied to other classes of bugs, i.e., what is next?  This includes:
-input validation, logic problems, leaking secrets, performance degradation,
-denial-of-service, command (e.g., SQL) injection, and concurrency attacks.
+%% \subsection{Focus Areas}
 
-\subsection{Focus Areas}
+%% Discussion centered on new / unexplored ways in which to secure networking
+%% protocols.  One primary area that was discussed centered around
+%% applying artificial diversity to protocols,  with implementation
+%% including potentially taking advantage of the ambiguity of protocols
+%% (i.e., differences in how messages in different protocols are
+%% interpreted by endpoints and middleboxes).  Some group members also
+%% felt like static and dynamic security analysis (to include fuzzing) of
+%% network protocols was an area that needed more work.
 
-\begin{enumerate}
-	\item The next generation of vulnerabilities (e.g., logic flaws)
-	\item Need good tools for analysis of race conditions / concurrency attacks
-\end{enumerate}
+%% \begin{enumerate}
 
-\section{Network Layer}
+%%   \item Protocol diversity for surviving attacks
 
-\subsection{Focus Areas}
+%%   \item Raise the bar on an attacker by forcing them to attack all
+%%   channels to be successful
 
-Discussion centered on new / unexplored ways in which to secure networking
-protocols.  One primary area that was discussed centered around
-applying artificial diversity to protocols,  with implementation
-including potentially taking advantage of the ambiguity of protocols
-(i.e., differences in how messages in different protocols are
-interpreted by endpoints and middleboxes).  Some group members also
-felt like static and dynamic security analysis (to include fuzzing) of
-network protocols was an area that needed more work.
+%%   \item Protocol dissection is not a solved problem (i.e., how to
+%%   understand and analyze new and unknown protocols)
 
-\begin{enumerate}
+%%   \begin{enumerate}
+%%   	\item Profile traffic to application-specific communication for
+%%   	composing into system-specific configuration baseline
+%%   \end{enumerate}
 
-  \item Protocol diversity for surviving attacks
+%% \end{enumerate}
 
-  \item Raise the bar on an attacker by forcing them to attack all
-  channels to be successful
+%% \section{User Layer}
 
-  \item Protocol dissection is not a solved problem (i.e., how to
-  understand and analyze new and unknown protocols)
+%% Discussion centered on trust of sensors in a system.  For instance,
+%% what can be done about sensors that lie (e.g., because they have been
+%% compromised), and also how to force a hacker to own all the sensors in
+%% a system in order to truly take control.  Control theory also needs to
+%% integrate cyber security into its processes, including adding threat
+%% model descriptions for control system design, and clean-slate designs
+%% to improve resilience.
 
-  \begin{enumerate}
-  	\item Profile traffic to application-specific communication for
-  	composing into system-specific configuration baseline
-  \end{enumerate}
+%% \subsection{Focus Areas}
 
-\end{enumerate}
+%% \begin{enumerate}
+%% 	\item Control theory research to improve algorithms
+%% 	to be resilient against faults caused by cyber attacks
+%% 	\item Sensor correlation to survive owning of sensors
+%% 	\item Clean slate control theory design for resilience and
+%% 	fail-safe recovery (including graceful degradation and restart in the
+%% 	event of a cyber attack)
+%% \end{enumerate}
 
-\section{User Layer}
+\bibliographystyle{abbrv}
+\bibliography{papers}
 
-Discussion centered on trust of sensors in a system.  For instance,
-what can be done about sensors that lie (e.g., because they have been
-compromised), and also how to force a hacker to own all the sensors in
-a system in order to truly take control.  Control theory also needs to
-integrate cyber security into its processes, including adding threat
-model descriptions for control system design, and clean-slate designs
-to improve resilience.
-
-\subsection{Focus Areas}
-
-\begin{enumerate}
-	\item Control theory research to improve algorithms
-	to be resilient against faults caused by cyber attacks
-	\item Sensor correlation to survive owning of sensors
-	\item Clean slate control theory design for resilience and
-	fail-safe recovery (including graceful degradation and restart in the
-	event of a cyber attack)
-\end{enumerate}
 \end{document}

### 2016-05-19 paper.tex
Fixed bugs according to Chrisper (bugs in writing)

@@
 runtime monitoring~\cite{Waksman2010}. In this way, the hardware
-itself is is monitoring itself, rather than external analysis of
+itself is monitoring itself, rather than external analysis of
 side-channels. However, this has all the same drawbacks of changing
@@
 obfuscated by an attacker, whereas a vulnerability is accidentally
-introduced by a benign developer. Therefore, Trojan behavior is more
+introduced by a benign developer. Therefore, Trojan behavior is 
 more difficult to detect, because there is an active adversary who is
@@
 
-Another approach has been to include run-time monitoring into the
+Another approach has been to include runtime monitoring into the
 firmware, so that compromises can be detected~\cite{Cui2010a,
@@
 
-Security of the application lays has received significant attention
+Security of the application layer has received significant attention
 from the research community. Because of this, we have divided the
-research into attack surface reduction, prevent control-flow
+research into the following categories: attack surface reduction, prevent control-flow
 hijacking, logic and implementation flaws, mitigation and recovery,
@@
 Significant progress in this area started with the idea of
-control-flow integrity, essentially ensuring dynamically at run-time
+control-flow integrity, essentially ensuring dynamically at runtime
 that the control-flow of the application does not

### 2016-05-28 paper.tex
Protection against malicious code is added - still working on it

@@
 
+% Sai: should this be 'arranged' instead of 'arraigned'?%Faris: I think it should be 'arranged'.
 The layers are arraigned from lowest to highest layer. The lower layer
@@
 computing system, yet they have a smaller attacker surface. However,
-the lower levels can be attacked by the upper levels. Also, the upper
+the lower levels can be attacked by exploiting the upper levels. Also, the upper
 levels trust the abstractions provided by the lower levels. For
 instance, the application-layer visualization of an industrial control
-system trusts that the data reported by the hardware layer it
+system trusts that the data reported by the hardware layer is
 accurate. If an attacker is able to control and manipulate the
@@
 The interconnected nature of the layers means that the security of the
-entire layers must be a priority. However, at the same time, the
-security improvement of any layer will have an amplifying affect of
+entire layer must be a priority. However, at the same time, the
+security improvement of any layer will have an amplifying effect on
 the security of the whole system. For instance, improvement to the
@@
 Hardware underpins all of modern computing. While most software is
-constructed and written in such as way as to abstract the
-peculiarities of the underlying hardware, the software could not
+constructed and written in such a way as to abstract the
+peculiarities of the underlying hardware, the software cannot
 execute without hardware. In our abstracted model of the computing
@@
 system: CPUs, SoCs, or any other integrated circuit (IC) that performs
-computation of communication on behalf of the upper computing layers.
-This also include networking cards and other peripheral devices
+computation or communication on behalf of the upper computing layers.
+This also includes networking cards and other peripheral devices
 (Bluetooth or USB devices), which may have their own special purpose
@@
 
-A key difficult in analyzing the security concerns of hardware is
+A key difficulty in analyzing the security concerns of hardware is
 reverse engineering the workings of the hardware. Traditional
@@
 of Trojan logic. This is malicious logic that is introduced into the
-hardware chip by a malicious adversary. Trojan introduction can be
+% adversary, by definition is a malicious entity.
+hardware chip by an adversary. Trojan introduction can be
 done at several points of the hardware lifecycle:
@@
 
-\item Design. A malicious insider can incorporate a Trojan into the
+\item \textbf{Design}. A malicious insider can incorporate a Trojan into the
   design of the hardware.
 
-\item Third-party IP. Oftentimes, hardware designers will take
+\item \textbf{Third-party IP}. Oftentimes, hardware designers will take
   advantage of the functionality of third-party components, using them
@@
 
-\item Compilation of Design. When the design of the hardware is
+\item \textbf{Compilation of Design}. When the design of the hardware is
   compiled, but before it is sent for fabrication, the design could be
@@
 
-\item Manufacture. A malicious manufacturer can introduce Trojan logic
-  into hardware after the design is received but before the hardware
+\item \textbf{Manufacture}. A malicious manufacturer can introduce Trojan logic into hardware after the design is received but before the hardware
 is fabricated.
 
-\item Delivery. A malicious hardware component containing Trojan logic
-  can be swapped with a benign component after fabrication but before
+\item \textbf{Delivery}. A malicious hardware component containing Trojan logic can be swapped with a benign component after fabrication but before
 delivery and installation.
@@
 The malicious Trojan logic behavior comes in three different forms:
+% Sai: is this three or two??
 
@@
 
-\item Trigger-based. The malicious behavior or logic is only expressed
-  when certain conditions are met. Consider a Trojan logic in a CPU,
+\item \textbf{Trigger-based}. The malicious behavior or logic is only expressed when certain conditions are met. Consider a Trojan logic in a CPU,
   which changes the executing code to Ring 0 (with the most
@@
 
-\item Constant. The malicious behavior or logic is continually
+\item \textbf{Constant}. The malicious behavior or logic is continually
   executed and subtly changes the output of the hardware in a way that
@@
 Of particular practical interest is a description of \emph{actual
-  Trojan logic} discovered in a the ProASIC3 Flash FPGA military
+  Trojan logic} discovered in the ProASIC3 Flash FPGA military
 chip~\cite{Skorobogatov2012}. While in this case it was not proved
@@
 runtime monitoring~\cite{Waksman2010}. In this way, the hardware
-itself is monitoring itself, rather than external analysis of
+itself is monitoring itself, 
+% Sai: can we rephrase this^ ? sounds weird %Faris : Removing the first 'itself' may solve the problem.
+rather than external analysis of
 side-channels. However, this has all the same drawbacks of changing
@@
 A key problem with hardware security is non-destructive IC
-examination, identification, and authentication. How to ensure that
+examination, identification, and authentication - how to ensure that
 the chip that is designed is what is actually fabricated, with no
@@
 hardware. This can also include changing or altering the firmware
-layer in such as way that it is not possible to exercise the Trojan
+layer in such a way that it is not possible to exercise the Trojan
 logic, potentially using randomization techniques similar to ASLR.
@@
 
+%Note: I did not remove the previous versions of those paragraphs
+%(commented them) so feel free to check the previous versions and 
+%revert them if my changes have problems.
+
+%Firmware is the layer of software that sits and interacts directly
+%with the hardware. Typically, the firmware is controlled from the
+%upper layers by device drivers in the Operating System.
+
+%Faris:
 Firmware is the layer of software that sits and interacts directly
-with the hardware. Typically, the firmware is controlled from the
-upper layers by device drivers in the Operating System.
-
-Many different hardware contain firmware, for instance the boot BIOS
-on a motherboard is a form of firmware, embedded systems often contain
-specific firmware, hardware peripheral such as network cards contain
-firmware as well. We consider all such areas as firmware.
-
-The key traits of firmware is that, although it is binary code, it is
+with the hardware. Typically, the firmware is controlled by the
+device drivers in the upper layers of the Operating System.
+
+%Many different hardware contain firmware, for instance the boot BIOS
+%on a motherboard is a form of firmware, embedded systems often contain
+%specific firmware, hardware peripherals such as network cards contain
+%firmware as well. We consider all such areas as firmware.
+
+%Faris:
+Many different hardware contain firmware. For instance: the boot BIOS
+on a motherboard, embedded systems that contain
+specific firmware, hardware peripherals such as network cards. 
+We consider all such areas as firmware.
+
+%The key traits of firmware is that, although it is binary code, it is
+%operating/interacting directly with the hardware, so there are no
+%%OS-style abstraction layers such as virtual memory, well-defined
+%system calls, or hardware abstraction. Therefore, each firmware is
+%binary code which is specifically tailored to a specific hardware
+%platform.
+
+%Faris:
+Although firmware is binary code, it is
 operating/interacting directly with the hardware, so there are no
 OS-style abstraction layers such as virtual memory, well-defined
-system calls, or hardware abstraction. Therefore, each firmware is
-binary code with is specifically tailed to a specific hardware
-platform.
-
+system calls, or hardware abstraction which is the key trait of firmware. 
+Therefore, each firmware is specifically tailored to a specific hardware platform.
+
+%Due to the hardware-specific nature of firmware, \emph{reverse
+%	engineering} is critical to understanding the firmware. In addition,
+%reverse engineering is particularly challenging because source code is
+%frequently unavailable, and vendors' desire to protect their
+%intellectual property makes them actively obfuscate their
+%firmware. All this means that, extracting semantically useful
+%information from firmware is a key first step to performing any
+%security analysis of the firmware.
+
+%Faris:
 Due to the hardware-specific nature of firmware, \emph{reverse
-  engineering} is critical to understanding the firmware. In addition,
-reverse engineering is particularly difficult because source code is
-frequently unavailable, and, vendors desire to protect their
-intellectual property, therefore, they actively obfuscate their
-firmware. All this means that, extracting semantically useful
-information from firmware is a key first step to performing any
+  engineering} is critical to understand it. In addition,
+reverse engineering is particularly challenging because the source code is
+frequently unavailable, and vendors' desire to protect their
+intellectual property makes them actively obfuscate their
+firmware. This means, extracting semantically useful
+information from firmware is the first and key step to perform any
 security analysis of the firmware.
@@
 
+%Vulnerabilities in the firmware code allow an attacker to execute
+%arbitrary code with the permissions of the firmware. The class of
+%vulnerabilities are not unique to firmware, as they are the same class
+%of vulnerabilities that infect traditional binary software (buffer
+%overflows, logic flaws, etc.). However, firmware typically has full
+%control over the entire computing system, therefore vulnerabilities in
+%firmware are much more severe than those in the application layer. In
+%addition, if an attacker is able to persist their exploit --- because the
+%compromise occurs at such a low layer of the computing system --- the
+%attacker can persist \emph{even after reinstalling the operating
+%	system or application}.
+
+%Faris:
 Vulnerabilities in the firmware code allow an attacker to execute
@@
 firmware are much more severe than those in the application layer. In
-addition, if an attacker is able to persist their exploit, because the
-compromise occurs at such a low layer of the computing system, the
-attacker can persist \emph{even after reinstalling the operating
+addition, the attacker may remain in the system because the
+compromise occurs at such a low layer of the computing system so the
+exploit persists \emph{even after reinstalling the operating
 system or application}.
 
+%The attack surface for firmware varies depending on where it is used.
+%In the most general case, the attacks can come from the upper layers,
+%specifically the Operating System or Hypervisor layers, as these are
+%the layers of the computing stack that can directly communicate with the
+%firmware. However, other firmware can be attacked either directly
+%through the Hardware layer itself, or through the data being
+%transmitted through the firmware from the hardware. For instance, a
+%network card firmware could be attacked through network traffic that
+%the firmware processes.
+
+%Faris:
 The attack surface for firmware varies depending on where it is used.
-In the most general case, the attacks can come from the upper layers,
+In the general case, the attacks can come from the upper layers,
 specifically the Operating System or Hypervisor layers, as these are
-the layers of the computing stack that can directly communicate to the
+the layers of the computing stack that can directly communicate with the
 firmware. However, other firmware can be attacked either directly
-through the Hardware layer itself, or through the data being
+through the Hardware layer, or through the data being
 transmitted through the firmware from the hardware. For instance, a
@@
 
-Trojan capabilities in the firmware itself can allow a system to be
+%Trojan capabilities in the firmware itself can allow a system to be
+%remotely controlled or accessed. This is similar to a vulnerability,
+%in that an attacker is able to control the firmware, however the key
+%difference lies in the fact that Trojan behavior will be intentionally
+%obfuscated by an attacker, whereas a vulnerability is accidentally
+%introduced by a benign developer. Therefore, Trojan behavior is 
+%more difficult to detect, because there is an active adversary who is
+%intentionally obfuscating the behavior.
+
+%Faris:
+Trojan capabilities in the firmware can allow a system to be
 remotely controlled or accessed. This is similar to a vulnerability,
 in that an attacker is able to control the firmware, however the key
-difference lies in the fact that Trojan behavior will be intentionally
-obfuscated by an attacker, whereas a vulnerability is accidentally
+difference lies in the fact that Trojan behavior can be intentionally
+obfuscated by an attacker, whereas a vulnerability can accidentally
 introduced by a benign developer. Therefore, Trojan behavior is 
-more difficult to detect, because there is an active adversary who is
-intentionally obfuscating the behavior.
-
+difficult to detect.
+
+%Vulnerabilities in the firmware and Trojan capabilities are related
+%because both require \emph{understanding} the behavior and semantics
+%of the firmware code. This is why reverse engineering is such a
+%critical component of securing the threats against the firmware layer:
+%in order to detect either vulnerabilities or Trojan functionality, an
+%automated tool must first understand the semantics of the particular
+%firmware/hardware combination.
+
+%Faris:
 Vulnerabilities in the firmware and Trojan capabilities are related
@@
 of the firmware code. This is why reverse engineering is such a
-critical component of securing the threats against the firmware layer:
-in order to detect either vulnerabilities or Trojan functionality, an
-automated tool must first understand the semantics of the particular
+critical component of securing the firmware layer against the threats.
+In order to detect vulnerabilities or trojan functionality, an
+automated tool must understand the semantics of the particular
 firmware/hardware combination.
@@
 \subsection{Research}
-
+%Faris: Some fixes below. 
 Researchers have attempted to identify real-world attacks on firmware
@@
 
-Other approaches have looked at how to prevent the tampering of
-firmware~\cite{Adelstein2002, Zhou2007, Zhou2009}. This protects from
+Other approaches are about how to prevent the tampering of
+firmware~\cite{Adelstein2002, Zhou2007, Zhou2009}. These protect from
 the threat of tampering or altering the firmware, however it does not
@@
 
-Another approach has been to include runtime monitoring into the
+Another approach includes runtime monitoring into the
 firmware, so that compromises can be detected~\cite{Cui2010a,
@@
 
-Recent work has looked at the problem of updating the design of the
+Recent work has looked into the problem of updating the design of the
 firmware so that it has security features, such as memory handling and
@@
 
-Very few work has looked at how to reverse engineer and understand the
+Very few work has looked into how to reverse engineer and understand the
 firmware~\cite{Zaddach2014}. This problem is of particular interest,
@@
 
-Recent work has looked at how to detect vulnerabilities in
+Recent work has looked into how to detect vulnerabilities in
 firmware~\cite{Davidson2013, Costin2014, Shoshitaishvili2015}. One of
@@
 \subsection{Suggested Focus Areas}
-
+%Faris: Some fixes below.
 Of all the layers, firmware has received the least attention from the
@@
 
-The major issue with firmware vulnerabilities is how to decouple
+The major challenge with preventing firmware vulnerabilities is how to decouple
 firmware from the hardware, potentially using emulation of the
@@
 including analysis of memory layouts, code modules, even answering the
-simple question of what in a firmware binary is code and what is data.
+simple question of what is code in a firmware binary and what is data.
 
 Finally, even if we find vulnerabilities or Trojan logic in a
-firmware, there is still the unanswered questions of how to patch or
+firmware, there is still the unanswered question of how to patch or
 update the functionality of the firmware while still guaranteeing the
@@
 significant attention from the research community, uses a CAN bus to
-communicate. Modern motherboards us either a Northbound or Southbound
-bus to communicate between the CPU and the other peripheral devices,
-such as USB.
+communicate. Modern motherboards use either a Northbound or Southbound
+bus --- such as USB --- to communicate between the CPU and other peripheral devices.
 
@@
 If the bus and interconnect layer offers advanced functionality, such
-as the DMA example, then devices that an attacker is able to connect
+as the DMA example, then the devices that an attacker is able to connect
 to the bus could take advantage of this functionality. In the case of
@@
 A major area of research in the Bus and Interconnect layer is work
-that shows the danger of a bus and interconnect that allows malicious
+that shows the dangers of a bus and interconnect that allows malicious
 devices to either impersonate devices on the bus or leverage bus
@@
 attack~\cite{Wolf2012, Wojtczuk2011}. However, as these hardware
-functionality must be implemented, it's impact in legacy systems is
+functionalities must be implemented, it's impact in legacy systems is
 limited.
@@
 interconnect. There is also the need to support logging of the bus or
-interconnect activity. This will enable to detection of a malicious
+interconnect activity. This will enable detection of a malicious
 device on the bus or interconnect.
 
-Another promising area of research is into the automated filtering of
+Another promising area of research is the automated filtering of
 the bus communication, potentially allowing an administrator to define
@@
 lower layers of the computing stack, so that multiple Operating
-Systems, which each assume that the are the sole users of the
+Systems, which each assume that they are the sole users of the
 underlying hardware, can run concurrently on the same hardware.
@@
 System) and the \emph{guest} is the Operating System that is running
-in the hypervisor. A host can run multiple guest Operating Systems.
+on the hypervisor. A host can run multiple guest Operating Systems.
 
@@
 Randomization techniques attempt to alter the environment of the
-operating system in such as way that the system continues to function
+operating system in such a way that the system continues to function
 but the attacker's exploit fails~\cite{Lin2009, Giuffrida2012}.
@@
 
-As driver code executes in kernel space, and are a frequent source of
+% either driver code execute or change `and' to `is'
+As driver code executes in kernel space, and is a frequent source of
 vulnerabilities~\cite{Kadav2012}, researchers have looked into how to
@@
 We believe that, while significant research has been done on the
-Operating System layer, there still remains work to be done on this
+Operating System layer, there still remains work to be done in this
 area. In particular, as the Operating System is frequently in contact
@@
 scope, from compiled languages such as C or C++ to scripting languages
-such as Python and Ruby. In short, all types of computing application
+such as Python and Ruby. In short, all types of computing applications
 are included in the application layer.
@@
 Due to abstractions provided by the lower layers, applications
-typically have less privilege than the lower layers of the computing
+typically have lesser privilege than the lower layers of the computing
 stack. This is due to the abstractions provided to the application
@@
 aspect is that applications frequently use shared libraries or other
-middleware, which as also considered to be part of the application
+middleware, which are also considered to be part of the application
 layer. In this way, multiple different applications can all use the
@@
 
-Applications, by definition must be accessible by users in
+Applications, by definition, must be accessible by users in
 order to perform their useful computation. Due to this accessibility,
@@
 
-At a high level the threats to the application layers is an attacker
+At a high level the threat to the application layer is attackers being
 able to influence the system to perform computation of their choosing.
@@
 the application. The key difference from normal flaws or bugs in
-software is that vulnerabilities allow the program to deviate from the
-behavior. These vulnerabilities range the gamut of possible causes
+software is that vulnerabilities allow the program to deviate from the expected behavior. These vulnerabilities range the gamut of possible causes
 such as programming errors or configuration mistakes. They can be in
@@
 views vulnerabilities as a probability, and by reducing the attack
-surface, the odds of an attacking being able to exploit a
+surface, the odds of an attack being able to exploit a
 vulnerability successfully is reduced as well.
@@
 Several research works attempt to reduce the attack surface. Some
-approaches try to statically analyzing for deadcode, then restrict the
+approaches try to statically analyze for deadcode, then restrict the
 execution of the deadcode during runtime~\cite{Kurmus2011}. Other ways
@@
 
-Modern attackers, exploiting binary applications, such as browsers,
+Modern attackers, exploiting binary applications such as browsers,
 media players, or PDF readers, attempt to hijack the control-flow of
@@
 techniques~\cite{Checkoway2010, Buchanan2008, Kayaalp2012, Chen2012,
-  Goktas2014, Schwartz2011} and applied them to different
+  Goktas2014, Schwartz2011} and applied them to a different
 architecture~\cite{Davi2012}.
@@
   Hiser2012, Wartell2012} or detecting ROP
-attacks\cite{Polychronakis2011}. Metastudies looked at the
+attacks\cite{Polychronakis2011}. Metastudies have looked at the
 effectiveness of these defenses~\cite{Skowyra2013, Schuster2014}.
@@
 or other programming language-based flaws. Another kind of
-vulnerability class are called logic (or implementation) flaws. These
+vulnerability class is called logic (or implementation) flaws. These
 flaws occur when the logic of the application is faulty. Consider an
-ecommerce application that allows a coupon to be reapplied to an order
+e-commerce application that allows a coupon to be reapplied to an order
 multiple times, driving the price to zero. Or if a user is able to
@@
 
-Other classes such as preventing access control
-vulnerabilities~\cite{dalton09:nemesis} or statically detecting
-them~\cite{Sun2011}. Privilege escalation
-vulnerabilities~\cite{Monshizadeh2014}, data disclosure
-vulnerabilities~\cite{Muthukumaran2015}
-
-Parameter tampering vulnerabilities are another type of logic
-flaw~\cite{Bisht2010, Bisht2011}
-
-Execution After Redirect are another type of logic flaw
-vulnerabilities in web applications~\cite{Doupe2011, Payet2013}
-
-Single sign-on vulnerabilities~\cite{Zhou2014}.
+Other research has looked at preventing access control
+vulnerabilities~\cite{dalton09:nemesis}, or statically detecting
+them~\cite{Sun2011}.
+Some research has looked at other types of logic
+flaws, such as
+Privilege escalation vulnerabilities~\cite{Monshizadeh2014}, data disclosure
+vulnerabilities~\cite{Muthukumaran2015}, single sign-on vulnerabilities~\cite{Zhou2014}, and parameter tampering vulnerabilities~\cite{Bisht2010, Bisht2011}.
+
+Execution After Redirect is another type of logic flaw
+vulnerability in web applications~\cite{Doupe2011, Payet2013}.
 
@@
 application~\cite{tan08:autoises, Livshits2009a}. Other approaches dynamically
-execute the application to infer likely invariants, then static
-analysis attempts to find paths that violation those
+execute the application to infer likely invariants, then use static
+analysis to attempt to find paths that violate those
 invariants~\cite{felmetsger10:logic}. Further work tried to find missing
@@
 
+% citation needed here for black-box testing
 Recent work has looked at detecting logic flaw vulnerabilities in web
-applications in a black-box manner---that is, no knowledge of the
+applications in a black-box manner---that is, with no knowledge of the
 application's source code~\cite{Pellegrino2014}.
@@
 \subsubsection{Modeling and monitoring execution}
+This subarea of application layer security concentrates on modeling, monitoring, and analyzing the execution of an application to verify that the system is executing as expected, and to detect any malicious behavior.
+This can be seen as a testing strategy towards application security.
+
+There has been some research in finding new approaches of test generation for systems.
+One such approach is Concolic Testing~\cite{sen2007concolic}, which combines symbolic execution with concrete execution of the code being tested.
+
+Research has been made to automatically generate test inputs for Concolic testing (so that all execution paths can be tested) for major languages like C, C++, and Java~\cite{Garg2013, Jayaraman2009, Sen2005}.
+Other research in this area includes an approach to combine Random testing~\cite{bird1983automatic} with Concolic execution~\cite{Majumdar2007}, automated Concolic testing for smartphone applications~\cite{anand2012automated}, and using Concolic testing in a distributed, scalable way~\cite{kim2012scalable}.
+There has also been some work on discussing concolic testing tools and their limitations~\cite{Qu2011}.
+
+There has been significant research in the area of analysis of an application.
+Dynamic analysis~\cite{ball1999concept} analyzes the code at run time, and much research in this area has been into Taint-based tracking~\cite{Chang2008}, Anomaly Detection~\cite{Cova2007, li11:BLOCK, Li2012}, Context-sensitive sanitization for popular languages and templating languages~\cite{Saxena2011, Samuel2011}, and Binary Analysis~\cite{Song2008}.
+Other work has looked at automated attack generation using model checking~\cite{martin08:goal_directed_mc}, and classifying the available countermeasures for the C/C++ languages~\cite{Younan2012}.
+
+The other approach to analyze applications is Static Analysis~\cite{cousot1977abstract}, where the code is analyzed at compile time. There has been considerable research in this area, spanning across
 
 \subsubsection{Protection against malicious code}
+One of the challenging areas of application security is allowing arbitrary code execution without having to compromise security.
+There are times when applications need to run user provided code or host another program, for instance browsers running web applications or apps being uploaded to app stores.
+In order to run codes safely and making sure that they are not executing malicious code, there are number of protection mechanisms that can be used, one of which is Sandboxing.
+it is also known as Software Fault Isolation (SFI)~\cite{Wahbe1993}.
+the idea is to first run the code in an isolated environment called fault domain in the applications address space and then prevent the code from accessing outside of that fault domain.
+
+Early research is focused on rewriting unsafe instructions and providing interface for system calls to restrict program's access~\cite{Wahbe1993,Mccamant2006}.
+Native Client (NaCl)~\cite{Yee2009a} is a sandbox provided by Google than offers x86 untrusted Native code execution in Chrome browser using x86 segmentation.
+Therefore, web developers can leverage the performance of executing native code in a safe environment.
+Xax~\cite{Douceur2008} is another similar technology that is in form of a browser plug-in that provides an abstraction layer for OS system calls.
+other research extend this technology by covering ARM and x86-64~\cite{Zhao2011, Sehr2010}.
+others attempt to reduce the overhead of performing rewriting the binaries and memory isolation~\cite{Ford2008,Jana2011}.
+Sandboxing is also used for executing untrusted native code in Android applications~\cite{Sun2014,Afonso2016}.
+
+The other approach towards protecting against malicious code is to analyze the code and detect malicious behavior.
+Despite all the efforts that have been made to keep malicious apps off the Android market, malicious apps can still find their way into the app store by bypassing the Google Bouncer using different techniques.
+Bouncer uses sandboxing techniques to run apps in a confined environment, but apps can make it harder for the Bouncer to analyze using different techniques such as obfuscating, to complicate the reverse engineering process, or they can check IP's to detect whether or not it belongs to Google, if apps could detect that they are being analyzed, they can hold off on running their malicious code while being analyzed.
+A lot of research has been done to fingerprint the Android analyzers to show that evading detection is possible~\cite{Maier2014,Jing2014}.
+Other approaches are using analysis to detect logic bombs, that is, when the application changes its behavior under certain conditions~\cite{Fratantonio2016}.
 

### 2016-05-29 paper.tex
finished proof-reading background section

--- 2016-05-19 paper.tex
@@
 
+% Sai: should this be 'arranged' instead of 'arraigned'?
 The layers are arraigned from lowest to highest layer. The lower layer

### 2016-05-30 paper.tex
finished proof-reading till 3.1
finished proof-reading till 3.2, added comments

@@
 runtime monitoring~\cite{Waksman2010}. In this way, the hardware
-itself is monitoring itself, rather than external analysis of
+itself is monitoring itself, 
+% Sai: can we rephrase this^ ? sounds weird
+rather than external analysis of
 side-channels. However, this has all the same drawbacks of changing

### 2016-05-31 paper.tex
finished proof-reading till 4.1

@@
 on a motherboard is a form of firmware, embedded systems often contain
-specific firmware, hardware peripheral such as network cards contain
+specific firmware, hardware peripherals such as network cards contain
 firmware as well. We consider all such areas as firmware.
@@
 system calls, or hardware abstraction. Therefore, each firmware is
-binary code with is specifically tailed to a specific hardware
+binary code which is specifically tailored to a specific hardware
 platform.
@@
 reverse engineering is particularly difficult because source code is
-frequently unavailable, and, vendors desire to protect their
-intellectual property, therefore, they actively obfuscate their
+frequently unavailable, and vendors desire to protect their
+intellectual property makes them actively obfuscate their
 firmware. All this means that, extracting semantically useful
@@
 firmware are much more severe than those in the application layer. In
-addition, if an attacker is able to persist their exploit, because the
-compromise occurs at such a low layer of the computing system, the
+addition, if an attacker is able to persist their exploit --- because the
+compromise occurs at such a low layer of the computing system --- the
 attacker can persist \emph{even after reinstalling the operating
@@
 specifically the Operating System or Hypervisor layers, as these are
-the layers of the computing stack that can directly communicate to the
+the layers of the computing stack that can directly communicate with the
 firmware. However, other firmware can be attacked either directly

### 2016-05-31 paper.tex
finished proof-reading till 5.1

@@
   engineering} is critical to understanding the firmware. In addition,
-reverse engineering is particularly difficult because source code is
-frequently unavailable, and vendors desire to protect their
+reverse engineering is particularly challenging because source code is
+frequently unavailable, and vendors' desire to protect their
 intellectual property makes them actively obfuscate their
@@
 Finally, even if we find vulnerabilities or Trojan logic in a
-firmware, there is still the unanswered questions of how to patch or
+firmware, there is still the unanswered question of how to patch or
 update the functionality of the firmware while still guaranteeing the

### 2016-06-06 paper.tex
finished writing concolic execution section, static and dynamic analysis left

@@
 \subsubsection{Modeling and monitoring execution}
+This subarea of application layer security concentrates on modeling, monitoring, and analyzing the execution of an application to verify that the system is executing as expected, and to detect any malicious behavior.
+This can be seen as a testing strategy towards application security.
+
+There has been some research in finding new approaches of test generation for systems.
+One such approach is Concolic Testing~\cite{sen2007concolic}, which combines symbolic execution with concrete execution of the code being tested.
+
+Research has been made to automatically generate test inputs for Concolic testing (so that all execution paths can be tested) for major languages like C, C++, and Java~\cite{Garg2013, Jayaraman2009, Sen2005}.
+Other research in this area includes an approach to combine Random testing~\cite{bird1983automatic} with Concolic execution~\cite{Majumdar2007}, automated Concolic testing for smartphone applications~\cite{anand2012automated}, and using Concolic testing in a distributed, scalable way~\cite{kim2012scalable}.
+There has also been some work on discussing concolic testing tools and their limitations~\cite{Qu2011}.
 

### 2016-06-07 paper.tex
finished writing dynamic analysis section, started on static analysis added few more papers to the bibliography

@@
 
+There has been significant research in the area of analysis of an application.
+Dynamic analysis~\cite{ball1999concept} analyzes the code at run time, and much research in this area has been into Taint-based tracking~\cite{Chang2008}, Anomaly Detection~\cite{Cova2007, li11:BLOCK, Li2012}, Context-sensitive sanitization for popular languages and templating languages~\cite{Saxena2011, Samuel2011}, and Binary Analysis~\cite{Song2008}.
+Other work has looked at automated attack generation using model checking~\cite{martin08:goal_directed_mc}, and classifying the available countermeasures for the C/C++ languages~\cite{Younan2012}.
+
+The other approach to analyze applications is Static Analysis~\cite{cousot1977abstract}, where the code is analyzed at compile time. There has been considerable research in this area, spanning across
+
 \subsubsection{Protection against malicious code}

### 2016-06-07 paper.tex
Some fixes/edits in Firmware section

@@
 
-% Sai: should this be 'arranged' instead of 'arraigned'?
+% Sai: should this be 'arranged' instead of 'arraigned'?%Faris: I think it should be 'arranged'.
 The layers are arraigned from lowest to highest layer. The lower layer
@@
 itself is monitoring itself, 
-% Sai: can we rephrase this^ ? sounds weird
+% Sai: can we rephrase this^ ? sounds weird %Faris : Removing the first 'itself' may solve the problem.
 rather than external analysis of
@@
 
+%Note: I did not remove the previous versions of those paragraphs
+%(commented them) so feel free to check the previous versions and 
+%revert them if my changes have problems.
+
+%Firmware is the layer of software that sits and interacts directly
+%with the hardware. Typically, the firmware is controlled from the
+%upper layers by device drivers in the Operating System.
+
+%Faris:
 Firmware is the layer of software that sits and interacts directly
-with the hardware. Typically, the firmware is controlled from the
-upper layers by device drivers in the Operating System.
-
-Many different hardware contain firmware, for instance the boot BIOS
-on a motherboard is a form of firmware, embedded systems often contain
-specific firmware, hardware peripherals such as network cards contain
-firmware as well. We consider all such areas as firmware.
-
-The key traits of firmware is that, although it is binary code, it is
+with the hardware. Typically, the firmware is controlled by the
+device drivers in the upper layers of the Operating System.
+
+%Many different hardware contain firmware, for instance the boot BIOS
+%on a motherboard is a form of firmware, embedded systems often contain
+%specific firmware, hardware peripherals such as network cards contain
+%firmware as well. We consider all such areas as firmware.
+
+%Faris:
+Many different hardware contain firmware. For instance: the boot BIOS
+on a motherboard, embedded systems that contain
+specific firmware, hardware peripherals such as network cards. 
+We consider all such areas as firmware.
+
+%The key traits of firmware is that, although it is binary code, it is
+%operating/interacting directly with the hardware, so there are no
+%%OS-style abstraction layers such as virtual memory, well-defined
+%system calls, or hardware abstraction. Therefore, each firmware is
+%binary code which is specifically tailored to a specific hardware
+%platform.
+
+%Faris:
+Although firmware is binary code, it is
 operating/interacting directly with the hardware, so there are no
 OS-style abstraction layers such as virtual memory, well-defined
-system calls, or hardware abstraction. Therefore, each firmware is
-binary code which is specifically tailored to a specific hardware
-platform.
-
+system calls, or hardware abstraction which is the key trait of firmware. 
+Therefore, each firmware is specifically tailored to a specific hardware platform.
+
+%Due to the hardware-specific nature of firmware, \emph{reverse
+%	engineering} is critical to understanding the firmware. In addition,
+%reverse engineering is particularly challenging because source code is
+%frequently unavailable, and vendors' desire to protect their
+%intellectual property makes them actively obfuscate their
+%firmware. All this means that, extracting semantically useful
+%information from firmware is a key first step to performing any
+%security analysis of the firmware.
+
+%Faris:
 Due to the hardware-specific nature of firmware, \emph{reverse
-  engineering} is critical to understanding the firmware. In addition,
-reverse engineering is particularly challenging because source code is
+  engineering} is critical to understand it. In addition,
+reverse engineering is particularly challenging because the source code is
 frequently unavailable, and vendors' desire to protect their
 intellectual property makes them actively obfuscate their
-firmware. All this means that, extracting semantically useful
-information from firmware is a key first step to performing any
+firmware. This means, extracting semantically useful
+information from firmware is the first and key step to perform any
 security analysis of the firmware.
@@
 firmware are much more severe than those in the application layer. In
-addition, if an attacker is able to persist their exploit --- because the
-compromise occurs at such a low layer of the computing system --- the
-attacker can persist \emph{even after reinstalling the operating
+addition, the attacker may remain in the system because the
+compromise occurs at such a low layer of the computing system so the
+exploit persists \emph{even after reinstalling the operating
 system or application}.
 
+%The attack surface for firmware varies depending on where it is used.
+%In the most general case, the attacks can come from the upper layers,
+%specifically the Operating System or Hypervisor layers, as these are
+%the layers of the computing stack that can directly communicate with the
+%firmware. However, other firmware can be attacked either directly
+%through the Hardware layer itself, or through the data being
+%transmitted through the firmware from the hardware. For instance, a
+%network card firmware could be attacked through network traffic that
+%the firmware processes.
+
+%Faris:
 The attack surface for firmware varies depending on where it is used.
-In the most general case, the attacks can come from the upper layers,
+In the general case, the attacks can come from the upper layers,
 specifically the Operating System or Hypervisor layers, as these are
@@
 firmware. However, other firmware can be attacked either directly
-through the Hardware layer itself, or through the data being
+through the Hardware layer, or through the data being
 transmitted through the firmware from the hardware. For instance, a
@@
 including analysis of memory layouts, code modules, even answering the
-simple question of what in a firmware binary is code and what is data.
+simple question of what is code in a firmware binary and what is data.
 

### 2016-06-11 paper.tex
Section 8.2.6 is completed

@@
 \subsubsection{Protection against malicious code}
+One of the challenging areas of application security is allowing arbitrary code execution without having to compromise security.
+There are situations in which applications need to run user provided code, reuse codes that were written in a non-type-safe language or host another program.
+For instance, browsers running web applications or apps being uploaded to app stores.
+In order to run codes safely and making sure that they are not executing malicious code, there are number of protection mechanisms that can be used, one of which is Sandboxing.
+it is also known as Software Fault Isolation (SFI)~\cite{Wahbe1993}.
+The idea is to run the code in an isolated environment within the applications address space called fault domain and then prevent the code from accessing outside of that fault domain.
+
+Early research is focused on rewriting unsafe instructions and providing interface for system calls to restrict program's access~\cite{Wahbe1993,Mccamant2006}.
+Native Client (NaCl)~\cite{Yee2009a} is a sandbox provided by Google that offers x86 untrusted Native code execution in Chrome browser using x86 memory segmentation.
+Therefore, web developers can leverage the performance of executing native code in a safe environment.
+Xax~\cite{Douceur2008} is another similar technology that is in form of a browser plug-in that provides an abstraction layer for OS system calls.
+Other research extends this technology by covering ARM and x86-64~\cite{Zhao2011, Sehr2010,Ansel}.
+Others attempt to reduce the overhead of performing binary rewriting and memory isolation~\cite{Ford2008,Jana2011}.
+Sandboxing is also used for executing untrusted native code in Android applications~\cite{Sun2014,Afonso2016}.
+
+The other approach towards protecting against malicious code is to analyze the code and detect malicious behavior~\cite{Zhou2012}.
+Despite all the efforts that have been made to keep malicious apps off the Android market, malicious apps can still find their way into the app store by bypassing the Google Bouncer~\cite{Lockheimer:2012} using different techniques.
+Bouncer uses sandboxing techniques to run apps in a confined environment, but researchers~\cite{oberheide2012dissecting} have shown that apps can fingerprint Bouncer and the environment in which they are being analyzed, using evasion and obfuscating techniques to complicate the reverse engineering process.
+If apps could detect that they are being analyzed, they can hold off on running their malicious code while being analyzed.
+A lot of research has been done to fingerprint the Android analyzers to show that evading detection is possible~\cite{Maier2014,Jing2014,Vidas2014}.
+Other approaches are using analysis to detect logic bombs, that is, when an application changes its behavior under certain conditions~\cite{Fratantonio2016}.
 

### 2016-06-11 fig_layers.tex
Computing Stack layers diagram is added

\begin{figure}
\centering
\begin{tikzpicture}[scale=0.50,node distance=3pt,outer sep=0pt,
platformlayer/.style={
  draw=darkgray,
  fill=nodecolor,
  rounded corners,
  text width=4.5cm,
  font={\sffamily\bfseries\color{black}},
  align=center,
  text height=9pt,
  text depth=6pt},
]
\node[platformlayer] (APP) {Application};
\node[platformlayer,below=of APP](OS) {Operating System};
\node[platformlayer,below=of OS] (HYP) {Hypervisor};
\node[platformlayer,below=of HYP](BUS){BUS and Interconnect};
\node[platformlayer,below=of BUS](FW) {Firmware};
\node[platformlayer,below=of FW](HW) {Hardware};

\begin{pgfonlayer}{background}
\draw[platformlayer,draw=black,fill=nodecolor!40]([xshift=-\framesep,yshift=1\framesep]current bounding box.north west)
rectangle ([xshift=\framesep,yshift=-\framesep]current bounding box.south east) 
node[below] at (0,2) {};

\end{pgfonlayer}
\end{tikzpicture}
\caption{Computing Platform Layers}
\label{fig:layers}
\end{figure}

### 2016-06-11 paper.tex
Computing Stack layers diagram is added

@@
 
+\usepackage{tikz}
+\usetikzlibrary{positioning}
+\definecolor{nodecolor}{RGB}{202,202,202}
+\newlength\framesep
+\setlength\framesep{16pt}
+\pgfdeclarelayer{background}
+\pgfsetlayers{background,main}
+\pgfkeys{
+	/tikz/node distance/.append code={
+		\pgfkeyssetvalue{/tikz/node distance value}{#1}
+	}
+}
+
 \widowpenalty=10000
@@
 hardware, firmware, bus, hypervisor, operating system, application,
-and network layers.
+and network layers (Figure~1).
 
@@
 that layer, and finally we will determine areas that require further
-research investment to ensure the security of the layer.
+research investment to ensure the security of the layer.\input{fig_layers}
 

### 2016-06-23 paper.tex
finished proof reading 8.2.6, great job yeganeh

@@
 This subarea of application layer security concentrates on modeling, monitoring, and analyzing the execution of an application to verify that the system is executing as expected, and to detect any malicious behavior.
-This can be seen as a testing strategy towards application security.
+This can be seen as a testing strategy toward application security.
 
@@
 Dynamic analysis~\cite{ball1999concept} analyzes the code at run time, and much research in this area has been into Taint-based tracking~\cite{Chang2008}, Anomaly Detection~\cite{Cova2007, li11:BLOCK, Li2012}, Context-sensitive sanitization for popular languages and templating languages~\cite{Saxena2011, Samuel2011}, and Binary Analysis~\cite{Song2008}.
-Other work has looked at automated attack generation using model checking~\cite{martin08:goal_directed_mc}, and classifying the available countermeasures for the C/C++ languages~\cite{Younan2012}.
+Other work has looked at automated attack generation using model checking~\cite{martin08:goal_directed_mc}, and classifying the available countermeasures for the C and C++ languages~\cite{Younan2012}.
 
@@
 \subsubsection{Protection against malicious code}
-One of the challenging areas of application security is allowing arbitrary code execution without having to compromise security.
-There are situations in which applications need to run user provided code, reuse codes that were written in a non-type-safe language or host another program.
-For instance, browsers running web applications or apps being uploaded to app stores.
-In order to run codes safely and making sure that they are not executing malicious code, there are number of protection mechanisms that can be used, one of which is Sandboxing.
-it is also known as Software Fault Isolation (SFI)~\cite{Wahbe1993}.
-The idea is to run the code in an isolated environment within the applications address space called fault domain and then prevent the code from accessing outside of that fault domain.
-
-Early research is focused on rewriting unsafe instructions and providing interface for system calls to restrict program's access~\cite{Wahbe1993,Mccamant2006}.
+One of the more challenging areas of application security is allowing arbitrary code execution without compromising security.
+There are situations in which applications need to run user provided code, reuse code that was written in a non-type-safe language, or host another program.
+For instance, browsers running web applications, or apps being uploaded to app stores.
+To run code safely and make sure that they are not executing malicious code, there are a number of protection mechanisms that can be used, one of which is Sandboxing, also known as Software Fault Isolation (SFI)~\cite{Wahbe1993}.
+The idea is to run the code in an isolated environment within the application's address space (called fault domain) and then prevent the code from accessing addresses outside that fault domain.
+
+Early research was focused on rewriting unsafe instructions and providing interfaces for system calls to restrict program's access~\cite{Wahbe1993, Mccamant2006}.
 Native Client (NaCl)~\cite{Yee2009a} is a sandbox provided by Google that offers x86 untrusted Native code execution in Chrome browser using x86 memory segmentation.
@@
 Xax~\cite{Douceur2008} is another similar technology that is in form of a browser plug-in that provides an abstraction layer for OS system calls.
-Other research extends this technology by covering ARM and x86-64~\cite{Zhao2011, Sehr2010,Ansel}.
-Others attempt to reduce the overhead of performing binary rewriting and memory isolation~\cite{Ford2008,Jana2011}.
-Sandboxing is also used for executing untrusted native code in Android applications~\cite{Sun2014,Afonso2016}.
-
-The other approach towards protecting against malicious code is to analyze the code and detect malicious behavior~\cite{Zhou2012}.
+Other research extends this technology by covering ARM and x86-64~\cite{Zhao2011, Sehr2010, Ansel}.
+Others attempt to reduce the overhead of performing binary rewriting and memory isolation~\cite{Ford2008, Jana2011}.
+Sandboxing is also used for executing untrusted native code in Android applications~\cite{Sun2014, Afonso2016}.
+
+The other approach toward protecting against malicious code is to analyze the code and detect malicious behavior~\cite{Zhou2012}.
 Despite all the efforts that have been made to keep malicious apps off the Android market, malicious apps can still find their way into the app store by bypassing the Google Bouncer~\cite{Lockheimer:2012} using different techniques.
@@
 If apps could detect that they are being analyzed, they can hold off on running their malicious code while being analyzed.
-A lot of research has been done to fingerprint the Android analyzers to show that evading detection is possible~\cite{Maier2014,Jing2014,Vidas2014}.
-Other approaches are using analysis to detect logic bombs, that is, when an application changes its behavior under certain conditions~\cite{Fratantonio2016}.
+A lot of research has been done to fingerprint the Android analyzers to show that evading detection is possible~\cite{Maier2014, Jing2014, Vidas2014}.
+Other approaches use analysis to detect logic bombs, i.e., when an application changes its behavior under certain conditions~\cite{Fratantonio2016}.
 

### 2016-06-25 paper.tex
fix(biblio): removed duplicates and updated references to be consistent

@@
 the rootkit from altering the Operating System
-kernel~\cite{Riley2008a, Wang2009, Grace2010, Gadaleta2011}.
+kernel~\cite{Riley2008, Wang2009, Grace2010, Gadaleta2011}.
 
@@
 There has been significant research in the area of analysis of an application.
-Dynamic analysis~\cite{ball1999concept} analyzes the code at run time, and much research in this area has been into Taint-based tracking~\cite{Chang2008}, Anomaly Detection~\cite{Cova2007, li11:BLOCK, Li2012}, Context-sensitive sanitization for popular languages and templating languages~\cite{Saxena2011, Samuel2011}, and Binary Analysis~\cite{Song2008}.
+Dynamic analysis~\cite{ball1999concept} analyzes the code at run time, and much research in this area has been into Taint-based tracking~\cite{Chang2008}, Anomaly Detection~\cite{Cova2007, li11:BLOCK, Li2012}, Context-sensitive sanitization for popular languages and templating languages~\cite{Saxena2011, Samuel2011}, Binary Analysis~\cite{Song2008}, and Logic vulnerabilities~{\cite{felmetsger10:logic}}.
 Other work has looked at automated attack generation using model checking~\cite{martin08:goal_directed_mc}, and classifying the available countermeasures for the C and C++ languages~\cite{Younan2012}.

### 2016-06-25 paper.tex
make all references lowercase, remove unnecessary tabs in references

--- 2016-06-23 paper.tex
@@
 Most research into hardware Trojans deal with using side-channel
-analysis to identify the malicious Trojan behavior~\cite{Lee2005,
-  Agrawal2007, Banga2008, Banga2008a, Jin2008, Du2010, Narasimhan2011,
-  Narasimhan2012, Cao2013, Davoodi2013, Dworak2013, Forte2013,
-  Liu2013, Liu2014, Bao2015}. This research uses power analysis, heat
+analysis to identify the malicious Trojan behavior~\cite{lee2005,
+  agrawal2007, banga2008, banga2008a, jin2008, du2010, narasimhan2011,
+  narasimhan2012, cao2013, davoodi2013, dworak2013, forte2013,
+  liu2013, liu2014, bao2015}. This research uses power analysis, heat
 analysis, or other side-effects to determine that the hardware is
@@
 Other research has demonstrated the attacks that are possible by
-actually implementing Trojan behavior~\cite{King2008, Jin2010,
-  Sturton2011, Liu2013, Jin2014, Zhang2014}. This line of research
+actually implementing Trojan behavior~\cite{king2008, jin2010,
+  sturton2011, liu2013, jin2014, zhang2014}. This line of research
 shows that Trojan logic is feasible for an attack.
@@
   Trojan logic} discovered in the ProASIC3 Flash FPGA military
-chip~\cite{Skorobogatov2012}. While in this case it was not proved
+chip~\cite{skorobogatov2012}. While in this case it was not proved
 that the logic introduced to the chip was intentionally malicious,
@@
 in order to improve the testability of the chip or to increase the
-effectiveness of the side-channel analysis~\cite{Chakraborty2008,
-  Li2008, Salmani2009, Hicks2010, Rosenfeld2011, Waksman2011,
-  Bhunia2013, Davoodi2013, Rajendran2013, Rolt2014}. This line of
+effectiveness of the side-channel analysis~\cite{chakraborty2008,
+  li2008, salmani2009, hicks2010, rosenfeld2011, waksman2011,
+  bhunia2013, davoodi2013, rajendran2013, rolt2014}. This line of
 research attempts to rethink designing the hardware itself. However, a
@@
 has investigated the problem of securing JTAG access to
-hardware~\cite{Rosenfeld2010, Ren2015}. The main idea is to allow only
+hardware~\cite{rosenfeld2010, ren2015}. The main idea is to allow only
 authenticated users to have access to the JTAG functionality, however
@@
 Another approach is to change the design of the chip to support
-runtime monitoring~\cite{Waksman2010}. In this way, the hardware
-itself is monitoring itself, 
-% Sai: can we rephrase this^ ? sounds weird %Faris : Removing the first 'itself' may solve the problem.
+runtime monitoring~\cite{waksman2010}. In this way, the hardware
+is monitoring itself, 
 rather than external analysis of
@@
 Recent approaches suggest to use machine-learning techniques to
-identify the Trojan logic~\cite{Haider2015}. Another approach attempts
+identify the Trojan logic~\cite{haider2015}. Another approach attempts
 to use a formal verification technique in order to prove that no
-malicious logic was added to the hardware~\cite{Guo2015}.
+malicious logic was added to the hardware~\cite{guo2015}.
 
 Finally, there are several surveys conducted on hardware security,
-which provide a good overview of the field~\cite{Wang2008,
-  Tehranipoor2010, Guin2014, Guin2014a, Guin2014b}.
+which provide a good overview of the field~\cite{wang2008,
+  tehranipoor2010, guin2014, guin2014a, guin2014b}.
 
@@
 Researchers have attempted to identify real-world attacks on firmware
-code~\cite{Tsow2006, Cui2010, Duflot2011, Basnight2013, Cui2013,
-  Zaddach2013, Zaddach2013a, Maskiewicz2014, Stuttgen2015}. These
+code~\cite{tsow2006, cui2010, duflot2011, basnight2013, cui2013,
+  zaddach2013, zaddach2013a, maskiewicz2014, stuttgen2015}. These
 attacks are particularly effective, and researchers have shown that
@@
 Other approaches are about how to prevent the tampering of
-firmware~\cite{Adelstein2002, Zhou2007, Zhou2009}. These protect from
+firmware~\cite{adelstein2002, zhou2007, zhou2009}. These protect from
 the threat of tampering or altering the firmware, however it does not
@@
 Another approach includes runtime monitoring into the
-firmware, so that compromises can be detected~\cite{Cui2010a,
-  Cui2011}. This approach is further complicated by the lack of
+firmware, so that compromises can be detected~\cite{cui2010a,
+  cui2011}. This approach is further complicated by the lack of
 abstractions in the firmware environment.
@@
 firmware so that it has security features, such as memory handling and
-avoiding information leakage~\cite{Koeberl2014, Kauer2007}. However,
+avoiding information leakage~\cite{koeberl2014, kauer2007}. However,
 forcing firmware developers to use such features is an open issue.
@@
 Very few work has looked into how to reverse engineer and understand the
-firmware~\cite{Zaddach2014}. This problem is of particular interest,
+firmware~\cite{zaddach2014}. This problem is of particular interest,
 as understanding the semantics of the firmware and how it interacts
@@
 Recent work has looked into how to detect vulnerabilities in
-firmware~\cite{Davidson2013, Costin2014, Shoshitaishvili2015}. One of
+firmware~\cite{davidson2013, costin2014, shoshitaishvili2015}. One of
 these has an excellent website\footnote{\url{http://firmware.re}}
 that contains the results of their reverse engineering and
-vulnerability analysis~\cite{Costin2014}.
+vulnerability analysis~\cite{costin2014}.
 
@@
 devices to either impersonate devices on the bus or leverage bus
-capabilities, particularly in the automotive area~\cite{Bonkoski2013,
-  Checkoway2011, Hoppe2008, Koscher2010, Rouf2010, Sang2010}.
+capabilities, particularly in the automotive area~\cite{bonkoski2013,
+  checkoway2011, hoppe2008, koscher2010, rouf2010, sang2010}.
 
 The other major area of research is to apply encryption to devices on
-the bus and interconnect~\cite{Groza2012, Herrewege2011, Jiang2012,
-  Schweppe2011, Stewin2014, Szilagyi2010}. In this way, malicious
+the bus and interconnect~\cite{groza2012, herrewege2011, jiang2012,
+  schweppe2011, stewin2014, szilagyi2010}. In this way, malicious
 devices on the bus cannot impersonate other devices. However, a key
@@
 attached to the bus and interconnect using attestation
-techniques~\cite{Li2010, Li2011}. These techniques attempt to ensure
+techniques~\cite{li2010, li2011}. These techniques attempt to ensure
 that only authorized and unmodified devices are attached to the bus
@@
 A key problem is how to perform secure updates to devices attached to
-the bus and the interconnect~\cite{Larson2008}. This is a challenge
+the bus and the interconnect~\cite{larson2008}. This is a challenge
 that also has similarities with the firmware layer.
@@
 between devices connect to the bus and interconnect
-layers~\cite{Schweppe2012}. Using these techniques, the researchers
+layers~\cite{schweppe2012}. Using these techniques, the researchers
 are able to identify malicious flows or potential vulnerabilities by
@@
 Other researchers have explored modeling either the system itself
-or the behavior of the devices on the bus~\cite{Stewin2013a,
-  Drolia2011}. If the device behavior or the system behavior does not
+or the behavior of the devices on the bus~\cite{stewin2013a,
+  drolia2011}. If the device behavior or the system behavior does not
 match the learned model, then an alert is triggered. In this way,
@@
 the functionality of the bus and interconnect
-layer~\cite{Vasudevan2012, Zhang2013, Zhou2009}. In this way, if the
+layer~\cite{vasudevan2012, zhang2013, zhou2009}. In this way, if the
 functionality is not there on the bus, then a malicious device cannot
@@
 bus and interconnect system to reduce the impact of an
-attack~\cite{Wolf2012, Wojtczuk2011}. However, as these hardware
+attack~\cite{wolf2012, wojtczuk2011}. However, as these hardware
 functionalities must be implemented, it's impact in legacy systems is
@@
 Finally, other research in this area has performed surveys of the
-research on this area~\cite{Kleberger2011, Studnia2013, Thom2008,
-  Wolf2004, Wolf2007, Wright2011, Zhao2002}.
+research on this area~\cite{kleberger2011, studnia2013, thom2008,
+  wolf2004, wolf2007, wright2011, zhao2002}.
 
@@
 Hypervisor to monitor the guest for security
-violations~\cite{Karger1991, Barham2003}, intrusion
-detections~\cite{Dunlap2002, Garfinkel2003a, Kourai2005, Chen2008,
-  Garfinkel2003b}, checking the integrity of the guest Operating
-system~\cite{Seshadri2006, McCune2010}, and identifying rootkits present
-in the guest Operating System~\cite{Jones2008, Grace2010}. This work
+violations~\cite{karger1991, barham2003}, intrusion
+detections~\cite{dunlap2002, garfinkel2003a, kourai2005, chen2008,
+  garfinkel2003b}, checking the integrity of the guest Operating
+system~\cite{seshadri2006, mccune2010}, and identifying rootkits present
+in the guest Operating System~\cite{jones2008, grace2010}. This work
 assumes that the guest can be compromised and that the Hypervisor is
@@
 guest-specific Operating System constructs. Therefore, research has
-investigated how to overcome this semantic gap~\cite{Jones2006,
-  Christodorescu2009, Srinivasan2011}.
+investigated how to overcome this semantic gap~\cite{jones2006,
+  christodorescu2009, srinivasan2011}.
 
 Further work in this line has approached the problem of how to monitor
-the hypervisor itself by leveraging hardware features~\cite{Azab2010}.
+the hypervisor itself by leveraging hardware features~\cite{azab2010}.
 
@@
 establishing a Trusted Computing Base, which validates the integrity
-of the lower levels for the guest Operating System~\cite{Arbaugh1997,
-  Sailer2004, Singaravelu2006} is important for ensuring the security
+of the lower levels for the guest Operating System~\cite{arbaugh1997,
+  sailer2004, singaravelu2006} is important for ensuring the security
 of the Hypervisor.
@@
 computing platform. Therefore, attempts have been made to reduce the
-trusted base of the Hypervisor itself~\cite{Murray2008, Shinagawa2009,
-  Colp2011, Wang2012}.
+trusted base of the Hypervisor itself~\cite{murray2008, shinagawa2009,
+  colp2011, wang2012}.
 
 Other research has looked at creating a secure
-Hypervisor~\cite{Robin2000, Belay2012, Szefer2011, Vasudevan2013} with
+Hypervisor~\cite{robin2000, belay2012, szefer2011, vasudevan2013} with
 the possibility of adding functionality to the Hypervisor to support
-closed guest Operating Systems~\cite{Garfinkel2003}, offering
+closed guest Operating Systems~\cite{garfinkel2003}, offering
 applications running in the guest Operating System to hide information
-from the guest Operating System~\cite{Chen2008, Yang2008, Jin2011,
-  Szefer2012, Azab2011, Zhang2011, Jin2011a, Szefer2011a}, secure
+from the guest Operating System~\cite{chen2008, yang2008, jin2011,
+  szefer2012, azab2011, zhang2011, jin2011a, szefer2011a}, secure
 information flow between the guest Operating System drivers and I/O
-devices~\cite{Cheng2013}, or designing the hypervisor so that it can
+devices~\cite{cheng2013}, or designing the hypervisor so that it can
 prove that it did not tamper with the guest Operating
-System~\cite{Gu2011}.
+System~\cite{gu2011}.
 
@@
 remove the duplicated components from the guest Operating System
-layer~\cite{Kivity2014, Madhavapeddy2015}, which effectively reduces
+layer~\cite{kivity2014, madhavapeddy2015}, which effectively reduces
 the attack surface of the Operating System layer.
@@
 boxes, such as firewalls and Intrusion Detection Systems, as a guest
-Operating System of a Hypervisor~\cite{Martins2014}.
+Operating System of a Hypervisor~\cite{martins2014}.
 
@@
 or prevent vulnerabilities, such as control-flow
-integrity~\cite{Wang2010} or manual vulnerability
+integrity~\cite{wang2010} or manual vulnerability
 analysis~\cite{Perez-Botero2013}.
@@
 which has looked at design flaws in Operating
-Systems~\cite{Attanasio1973}. Further research looked at reducing the
+Systems~\cite{attanasio1973}. Further research looked at reducing the
 permissions of the \emph{root} users, essentially compartmentalizing
-the privilege of the Operating System~\cite{Kamp2000, Kurmus2014a}.
+the privilege of the Operating System~\cite{kamp2000, kurmus2014a}.
 
@@
 important research topic in Operating System
-security~\cite{Petroni2004, Wang2005, Petroni2007, Payne2008,
-  Sharif2009, Yin2010, Bianchi2012}. Another approach is to prevent
+security~\cite{petroni2004, wang2005, petroni2007, payne2008,
+  sharif2009, yin2010, bianchi2012}. Another approach is to prevent
 the rootkit from altering the Operating System
-kernel~\cite{Riley2008a, Wang2009, Grace2010, Gadaleta2011}.
+kernel~\cite{riley2008, wang2009, grace2010, gadaleta2011}.
 
@@
 to use and exploit those vulnerabilities. Static analysis techniques
-have been used~\cite{Chou2001, Chen2011}.
+have been used~\cite{chou2001, chen2011}.
 
 Other vulnerability defenses include defenses for specific
-vulnerability classes, such as buffer overflows~\cite{Dalton2008}.
+vulnerability classes, such as buffer overflows~\cite{dalton2008}.
 Randomization techniques attempt to alter the environment of the
 operating system in such a way that the system continues to function
-but the attacker's exploit fails~\cite{Lin2009, Giuffrida2012}.
+but the attacker's exploit fails~\cite{lin2009, giuffrida2012}.
 Control-flow integrity techniques have been applied to Operating
-System code as well~\cite{Li2011a}.
+System code as well~\cite{li2011a}.
 
@@
 protect applications from a malicious or compromised Operating
-System~\cite{Ta-min2006, Onarlioglu2013, Li2014}.
+System~\cite{ta-min2006, onarlioglu2013, li2014}.
 
@@
 As driver code executes in kernel space, and is a frequent source of
-vulnerabilities~\cite{Kadav2012}, researchers have looked into how to
-make drivers more robust~\cite{Ganapathy2008, Hruby2012} or isolated
-from the rest of the kernel~\cite{Boyd-Wickizer2010}.
+vulnerabilities~\cite{kadav2012}, researchers have looked into how to
+make drivers more robust~\cite{ganapathy2008, hruby2012} or isolated
+from the rest of the kernel~\cite{boyd-wickizer2010}.
 
@@
 without injecting any additional code, using Return Oriented
-Programming techniques~\cite{Hund2009}.
+Programming techniques~\cite{hund2009}.
 
 A completely different approach is to formally prove the correctness
-of the Operating System kernel~\cite{Klein2009}.
+of the Operating System kernel~\cite{klein2009}.
 
@@
 researchers have explored how to apply an update to the Operating
-System without rebooting the system~\cite{Arnold2009}.
+System without rebooting the system~\cite{arnold2009}.
 
 Some work has been done on removing the operating system
-entirely~\cite{Madhavapeddy2010} or trimming/tailoring parts of the
-Operating System~\cite{Kurmus2011, Madhavapeddy2013, Howell2013,
-  Kurmus2013, Kurmus2014}. This has the security benefit of reducing
+entirely~\cite{madhavapeddy2010} or trimming/tailoring parts of the
+Operating System~\cite{kurmus2011, madhavapeddy2013, howell2013,
+  kurmus2013, kurmus2014}. This has the security benefit of reducing
 the attack surface of the Operating System.
@@
 Attempts have been made to measure an application's attack
-surface~\cite{Howard2003, Manadhata2007, Manadhata2008, Ruprecht2014}.
+surface~\cite{howard2003, manadhata2007, manadhata2008, ruprecht2014}.
 This has the benefit of attempting to quantify the attack surface of
@@
 approaches try to statically analyze for deadcode, then restrict the
-execution of the deadcode during runtime~\cite{Kurmus2011}. Other ways
+execution of the deadcode during runtime~\cite{kurmus2011}. Other ways
 attempt to leverage compile-time configurability to reduce the attack
-surface~\cite{Tartler2012, Kurmus2013, Ruprecht2014}. A different
+surface~\cite{tartler2012, kurmus2013, ruprecht2014}. A different
 approach is to reduce the attack surface by partitioning the
-application into different privilege levels~\cite{Barth2010, Xu2012}.
+application into different privilege levels~\cite{barth2010, xu2012}.
 Other approaches attempt to restrict the communication between
 applications, in order to reduce opportunities for the attacker to
-input data to the system~\cite{Kantola2012}.
+input data to the system~\cite{kantola2012}.
 
 Finally, other approaches attempt to quantify the attack surface and
-measure the actual reduction~\cite{Kurmus2014}.
+measure the actual reduction~\cite{kurmus2014}.
 
@@
 hijacking, specifically memory corruption vulnerabilities such as
-buffer overflows, heap overflows, and format strings~\cite{Szekeres2013}.
+buffer overflows, heap overflows, and format strings~\cite{szekeres2013}.
 
@@
 that the control-flow of the application does not
-deviate~\cite{Abadi2005, Abadi2009}. The first implementations were
+deviate~\cite{abadi2005, abadi2009}. The first implementations were
 deemed infeasible, so later research attempted to improve the
-efficiency of control-flow integrity~\cite{Zhang2013, Tice2014,
-  Bletsch2011, Zhang2013, Zhang2013a}
+efficiency of control-flow integrity~\cite{zhang2013, tice2014,
+  bletsch2011, zhang2013, zhang2013a}.
 
@@
 application, so that the attacker can hijack the control-flow of the
-application without introducing new code~\cite{Roemer2012,
-  Bletsch2010, Tran2011}. Researchers have also improved these ROP
-techniques~\cite{Checkoway2010, Buchanan2008, Kayaalp2012, Chen2012,
-  Goktas2014, Schwartz2011} and applied them to a different
-architecture~\cite{Davi2012}.
-
-Other research has looked at defending ROP attacks~\cite{Pappas2012,
-  Davi2009, Davi2011, Onarlioglu2010, Gupta2013, Pappas2012, Lu2011,
-  Lu2011a, Homescu2012, Gokta2014, Cheng2014, Pappas2013, Snow2013,
-  Hiser2012, Wartell2012} or detecting ROP
-attacks\cite{Polychronakis2011}. Metastudies have looked at the
-effectiveness of these defenses~\cite{Skowyra2013, Schuster2014}.
+application without introducing new code~\cite{roemer2012,
+  bletsch2010, tran2011}. Researchers have also improved these ROP
+techniques~\cite{checkoway2010, buchanan2008, kayaalp2012, chen2012,
+  goktas2014, schwartz2011} and applied them to a different
+architecture~\cite{davi2012}.
+
+Other research has looked at defending ROP attacks~\cite{pappas2012,
+  davi2009, davi2011, onarlioglu2010, gupta2013, pappas2012, lu2011,
+  lu2011a, homescu2012, gokta2014, cheng2014, pappas2013, snow2013,
+  hiser2012, wartell2012} or detecting ROP
+attacks\cite{polychronakis2011}. Metastudies have looked at the
+effectiveness of these defenses~\cite{skowyra2013, schuster2014}.
 
@@
 control-flow graph, an attacker can still hijack the control
-flow~\cite{Carlini2015, Schuster2015, Carlini2014, Davi2014}.
+flow~\cite{carlini2015, schuster2015, carlini2014, davi2014}.
 
@@
 work looked at detecting state violations in web
-applications~\cite{Cova2007}.
+applications~\cite{cova2007}.
 
@@
 vulnerabilities~\cite{dalton09:nemesis}, or statically detecting
-them~\cite{Sun2011}.
+them~\cite{sun2011}.
 Some research has looked at other types of logic
 flaws, such as
-Privilege escalation vulnerabilities~\cite{Monshizadeh2014}, data disclosure
-vulnerabilities~\cite{Muthukumaran2015}, single sign-on vulnerabilities~\cite{Zhou2014}, and parameter tampering vulnerabilities~\cite{Bisht2010, Bisht2011}.
+Privilege escalation vulnerabilities~\cite{monshizadeh2014}, data disclosure
+vulnerabilities~\cite{muthukumaran2015}, single sign-on vulnerabilities~\cite{zhou2014}, and parameter tampering vulnerabilities~\cite{bisht2010, bisht2011}.
 
 Execution After Redirect is another type of logic flaw
-vulnerability in web applications~\cite{Doupe2011, Payet2013}.
+vulnerability in web applications~\cite{doupe2011, payet2013}.
 
@@
 security (authorization) checks automatically, without prior knowledge
-of the checks~\cite{Son2011, Son2013}.
+of the checks~\cite{son2011, son2013}.
 
@@
 applications in a black-box manner---that is, with no knowledge of the
-application's source code~\cite{Pellegrino2014}.
+application's source code~\cite{pellegrino2014}.
 
@@
 
-Research has been made to automatically generate test inputs for Concolic testing (so that all execution paths can be tested) for major languages like C, C++, and Java~\cite{Garg2013, Jayaraman2009, Sen2005}.
-Other research in this area includes an approach to combine Random testing~\cite{bird1983automatic} with Concolic execution~\cite{Majumdar2007}, automated Concolic testing for smartphone applications~\cite{anand2012automated}, and using Concolic testing in a distributed, scalable way~\cite{kim2012scalable}.
-There has also been some work on discussing concolic testing tools and their limitations~\cite{Qu2011}.
+Research has been made to automatically generate test inputs for Concolic testing (so that all execution paths can be tested) for major languages like C, C++, and Java~\cite{garg2013, jayaraman2009, sen2005}.
+Other research in this area includes an approach to combine Random testing~\cite{bird1983automatic} with Concolic execution~\cite{majumdar2007}, automated Concolic testing for smartphone applications~\cite{anand2012automated}, and using Concolic testing in a distributed, scalable way~\cite{kim2012scalable}.
+There has also been some work on discussing concolic testing tools and their limitations~\cite{qu2011}.
 
 There has been significant research in the area of analysis of an application.
-Dynamic analysis~\cite{ball1999concept} analyzes the code at run time, and much research in this area has been into Taint-based tracking~\cite{Chang2008}, Anomaly Detection~\cite{Cova2007, li11:BLOCK, Li2012}, Context-sensitive sanitization for popular languages and templating languages~\cite{Saxena2011, Samuel2011}, and Binary Analysis~\cite{Song2008}.
-Other work has looked at automated attack generation using model checking~\cite{martin08:goal_directed_mc}, and classifying the available countermeasures for the C and C++ languages~\cite{Younan2012}.
+Dynamic analysis~\cite{ball1999concept} analyzes the code at run time, and much research in this area has been into Taint-based tracking~\cite{chang2008}, Anomaly Detection~\cite{cova2007, li11:block, li2012}, Context-sensitive sanitization for popular languages and templating languages~\cite{saxena2011, samuel2011}, Binary Analysis~\cite{song2008}, and Logic vulnerabilities~\cite{felmetsger2010}.
+Other work has looked at automated attack generation using model checking~\cite{martin08:goal_directed_mc}, and classifying the available countermeasures for the C and C++ languages~\cite{younan2012}.
 
@@
 For instance, browsers running web applications, or apps being uploaded to app stores.
-To run code safely and make sure that they are not executing malicious code, there are a number of protection mechanisms that can be used, one of which is Sandboxing, also known as Software Fault Isolation (SFI)~\cite{Wahbe1993}.
+To run code safely and make sure that they are not executing malicious code, there are a number of protection mechanisms that can be used, one of which is Sandboxing, also known as Software Fault Isolation (SFI)~\cite{wahbe1993}.
 The idea is to run the code in an isolated environment within the application's address space (called fault domain) and then prevent the code from accessing addresses outside that fault domain.
 
-Early research was focused on rewriting unsafe instructions and providing interfaces for system calls to restrict program's access~\cite{Wahbe1993, Mccamant2006}.
-Native Client (NaCl)~\cite{Yee2009a} is a sandbox provided by Google that offers x86 untrusted Native code execution in Chrome browser using x86 memory segmentation.
+Early research was focused on rewriting unsafe instructions and providing interfaces for system calls to restrict program's access~\cite{wahbe1993, mccamant2006}.
+Native Client (NaCl)~\cite{yee2009} is a sandbox provided by Google that offers x86 untrusted Native code execution in Chrome browser using x86 memory segmentation.
 Therefore, web developers can leverage the performance of executing native code in a safe environment.
-Xax~\cite{Douceur2008} is another similar technology that is in form of a browser plug-in that provides an abstraction layer for OS system calls.
-Other research extends this technology by covering ARM and x86-64~\cite{Zhao2011, Sehr2010, Ansel}.
-Others attempt to reduce the overhead of performing binary rewriting and memory isolation~\cite{Ford2008, Jana2011}.
-Sandboxing is also used for executing untrusted native code in Android applications~\cite{Sun2014, Afonso2016}.
-
-The other approach toward protecting against malicious code is to analyze the code and detect malicious behavior~\cite{Zhou2012}.
-Despite all the efforts that have been made to keep malicious apps off the Android market, malicious apps can still find their way into the app store by bypassing the Google Bouncer~\cite{Lockheimer:2012} using different techniques.
+Xax~\cite{douceur2008} is another similar technology that is in form of a browser plug-in that provides an abstraction layer for OS system calls.
+Other research extends this technology by covering ARM and x86-64~\cite{zhao2011, sehr2010, ansel}.
+Others attempt to reduce the overhead of performing binary rewriting and memory isolation~\cite{ford2008, jana2011}.
+Sandboxing is also used for executing untrusted native code in Android applications~\cite{sun2014, afonso2016}.
+
+The other approach toward protecting against malicious code is to analyze the code and detect malicious behavior~\cite{zhou2012}.
+Despite all the efforts that have been made to keep malicious apps off the Android market, malicious apps can still find their way into the app store by bypassing the Google Bouncer~\cite{lockheimer:2012} using different techniques.
 Bouncer uses sandboxing techniques to run apps in a confined environment, but researchers~\cite{oberheide2012dissecting} have shown that apps can fingerprint Bouncer and the environment in which they are being analyzed, using evasion and obfuscating techniques to complicate the reverse engineering process.
 If apps could detect that they are being analyzed, they can hold off on running their malicious code while being analyzed.
-A lot of research has been done to fingerprint the Android analyzers to show that evading detection is possible~\cite{Maier2014, Jing2014, Vidas2014}.
-Other approaches use analysis to detect logic bombs, i.e., when an application changes its behavior under certain conditions~\cite{Fratantonio2016}.
+A lot of research has been done to fingerprint the Android analyzers to show that evading detection is possible~\cite{maier2014, jing2014, vidas2014}.
+Other approaches use analysis to detect logic bombs, i.e., when an application changes its behavior under certain conditions~\cite{fratantonio2016}.
 

### 2016-06-25 paper.tex
adding todos, and adding a pre-commit-hook
mitigation & recovery part updated

@@
 
-% Sai: should this be 'arranged' instead of 'arraigned'?%Faris: I think it should be 'arranged'.
-The layers are arraigned from lowest to highest layer. The lower layer
+The layers are arranged from lowest to highest layer. The lower layer
 has more control over the computing system, while the upper layers
@@
 The malicious Trojan logic behavior comes in three different forms:
-% Sai: is this three or two??
+% TODO: Sai: is this three or two??
 
@@
 
+Mitigation and recovery is of major concern when it comes to application and systems security.
+This area of research is important because recovery of a tainted system is the first concern post attack or exploitation.
+
+Researchers have found some VM-based approaches in order to recover from intrusions.~\cite{zhang2009b} suggest one such approach.
+They try to achieve recovery in three phases, Routine execution phase, Damage assessment phase and Recovery phase.
+This framework tries to log pre-attack states and system calls and takes forward or reverse taint-tracking approach.
+Recovery phase tries to Quarantine the malicious files and objects identified during the Damage assessment.
+
+A similar VM-based approach is taken in~\cite{oliveria2008b}.
+The Bezoar framework proposed in this paper tries to recover a system based on monitoring of propagation bytes in the system after hijacking attack has been detected.
+The framework replays the checkpointed run while ignoring inputs from malicious sources.
+
+Automated intrusion recovery system is proposed in~\cite{goel2005b}.
+Researchers have designed automated resolution policies which determines set of tainted files.
+Taser recovery system~\cite{goel2005b}, is designed based on following models.
+Recovery model, Audito, Analyzer and Resolver.
+Techniques like naming conventions and file permissions, logging of all events, dependency of files, process and sockets are used by Taser in order to analyze intrusion.
+Similar to Recovery phase approach taken in~\cite{zhang2009b}, Taser uses the Resolver model proposed to go back to the state of the operating system before intrusion.
+
+The research in this area of mitigation and recovery is majorly done taking the VM-based approach in consideration.
+
 \subsubsection{Modeling and monitoring execution}

### 2016-06-25 paper.tex
proofread 8.2.4, left comments

@@
 
-Mitigation and recovery is of major concern when it comes to application and systems security.
-This area of research is important because recovery of a tainted system is the first concern post attack or exploitation.
-
+Mitigation and recovery is a major area of concern when it comes to application security. 
+% needs sentence about importance of mitigation
+This is because recovery of a tainted system is the first concern post attack or exploitation.
+
+%% TODO: this needs more stuff about mitigation, theres nothing about mitigation here.
+%% TODO: this para needs grammar-check: check usage of commas, colons, etc.
+%% TODO: in order to --> to, repetitive word `tries'
+%% TODO: expand VM at least first time of usage, abbreviate it after
 Researchers have found some VM-based approaches in order to recover from intrusions.~\cite{zhang2009b} suggest one such approach.
@@
 
-A similar VM-based approach is taken in~\cite{oliveria2008b}.
+%% TODO: this has passive voice, try to avoid it
+A similar VM-based approach is taken in~\cite{oliveira2008b}.
 The Bezoar framework proposed in this paper tries to recover a system based on monitoring of propagation bytes in the system after hijacking attack has been detected.
@@
 
+%% TODO: grammar-check
 Automated intrusion recovery system is proposed in~\cite{goel2005b}.

### 2016-07-31 paper.tex
Addressed To-Do's in mititgation and recovery section and added some content

--- 2016-06-25 paper.tex
@@
 \subsubsection{Mitigation and recovery}
-
+%% Additions and changes made.
 Mitigation and recovery is of major concern when it comes to application and systems security.
@@
 
-Researchers have found some VM-based approaches in order to recover from intrusions.~\cite{zhang2009b} suggest one such approach.
+Researchers have found some Virtual Machine based approaches in order to recover from intrusions.~\cite{zhang2009b} suggest one such approach.
 They try to achieve recovery in three phases, Routine execution phase, Damage assessment phase and Recovery phase.
@@
 A similar VM-based approach is taken in~\cite{oliveria2008b}.
-The Bezoar framework proposed in this paper tries to recover a system based on monitoring of propagation bytes in the system after hijacking attack has been detected.
+The Bezoar framework proposed in this paper recovers a system based on monitoring of propagation bytes in the system after hijacking attack has been detected.
 The framework replays the checkpointed run while ignoring inputs from malicious sources.
+Checkpointing methodolgy is also used in ExecRecorder[cite], which works on the idea of lod-based recovery.
+They propose a method to checkpoint system states and logging nondeterministic events.
+Recovery of the system is achieved by replaying the execution post attack and based on the log generated the system mitigates the attack execution during the reply.
+
+Recovering programs from software bugs is also a method proposed for mitigation of software failures.
+Rx~\cite{qin2007} thoroughly explains a method based on checkpoiting to detect and recover from software bugs.
+This framework detects software bugs in a program using sensors and once the bug is detected mitigation is done by changing the the execution environment dynamically.
+They have tabulated results displaying the mitigation of software bugs such as buffer overflow and stack overflow.
 
@@
 Researchers have designed automated resolution policies which determines set of tainted files.
-Taser recovery system~\cite{goel2005b}, is designed based on following models.
-Recovery model, Audito, Analyzer and Resolver.
+Taser recovery system~\cite{goel2005b} is designed based on following models.
+Recovery model, Auditor, Analyzer and Resolver.
 Techniques like naming conventions and file permissions, logging of all events, dependency of files, process and sockets are used by Taser in order to analyze intrusion.

### 2016-07-31 paper.tex

@@
 
+Mitigation and recovery is a major area of concern when it comes to application security. 
+% needs sentence about importance of mitigation
+This is because recovery of a tainted system is the first concern post attack or exploitation.
+
+%% TODO: this needs more stuff about mitigation, theres nothing about mitigation here.
+%% TODO: this para needs grammar-check: check usage of commas, colons, etc.
+%% TODO: in order to --> to, repetitive word `tries'
+%% TODO: expand VM at least first time of usage, abbreviate it after
+Researchers have found some VM-based approaches in order to recover from intrusions.~\cite{zhang2009b} suggest one such approach.
+They try to achieve recovery in three phases, Routine execution phase, Damage assessment phase and Recovery phase.
+This framework tries to log pre-attack states and system calls and takes forward or reverse taint-tracking approach.
+Recovery phase tries to Quarantine the malicious files and objects identified during the Damage assessment.
+
+%% TODO: this has passive voice, try to avoid it
+A similar VM-based approach is taken in~\cite{oliveira2008b}.
+The Bezoar framework proposed in this paper tries to recover a system based on monitoring of propagation bytes in the system after hijacking attack has been detected.
+The framework replays the checkpointed run while ignoring inputs from malicious sources.
+
+%% TODO: grammar-check
+Automated intrusion recovery system is proposed in~\cite{goel2005b}.
+Researchers have designed automated resolution policies which determines set of tainted files.
+Taser recovery system~\cite{goel2005b}, is designed based on following models.
+Recovery model, Audito, Analyzer and Resolver.
+Techniques like naming conventions and file permissions, logging of all events, dependency of files, process and sockets are used by Taser in order to analyze intrusion.
+Similar to Recovery phase approach taken in~\cite{zhang2009b}, Taser uses the Resolver model proposed to go back to the state of the operating system before intrusion.
+
+The research in this area of mitigation and recovery is majorly done taking the VM-based approach in consideration.
+
 \subsubsection{Modeling and monitoring execution}

### 2016-08-21 paper.tex
commiting changes
Grammatical corrections made

--- 2016-07-31 paper.tex
@@
 %% Additions and changes made.
-Mitigation and recovery is of major concern when it comes to application and systems security.
-This area of research is important because recovery of a tainted system is the first concern post attack or exploitation.
+%%Changes made after grammerly -- thanks Sai..its superb tool
+Mitigation and recovery are of major concern when it comes to application and systems security.
+This area of research is important because recovery of a tainted system is the first concern post-attack or exploitation.
 
@@
 The framework replays the checkpointed run while ignoring inputs from malicious sources.
-Checkpointing methodolgy is also used in ExecRecorder[cite], which works on the idea of lod-based recovery.
+Checkpointing methodology is also used in ExecRecorder[cite], which works on the idea of log-based recovery.
 They propose a method to checkpoint system states and logging nondeterministic events.
@@
 Recovering programs from software bugs is also a method proposed for mitigation of software failures.
-Rx~\cite{qin2007} thoroughly explains a method based on checkpoiting to detect and recover from software bugs.
-This framework detects software bugs in a program using sensors and once the bug is detected mitigation is done by changing the the execution environment dynamically.
+Rx~\cite{qin2007} thoroughly explains a method based on checkpointing to detect and recover from software bugs.
+This framework detects software bugs in a program using sensors and once the bug is detected mitigation is done by changing the execution environment dynamically.
 They have tabulated results displaying the mitigation of software bugs such as buffer overflow and stack overflow.
@@
 Automated intrusion recovery system is proposed in~\cite{goel2005b}.
-Researchers have designed automated resolution policies which determines set of tainted files.
+Researchers have designed automated resolution policies which determine set of tainted files.
 Taser recovery system~\cite{goel2005b} is designed based on following models.
 Recovery model, Auditor, Analyzer and Resolver.
-Techniques like naming conventions and file permissions, logging of all events, dependency of files, process and sockets are used by Taser in order to analyze intrusion.
-Similar to Recovery phase approach taken in~\cite{zhang2009b}, Taser uses the Resolver model proposed to go back to the state of the operating system before intrusion.
-
-The research in this area of mitigation and recovery is majorly done taking the VM-based approach in consideration.
+Techniques like naming conventions and file permissions, logging of all events, the dependency of files, process and sockets are used by Taser in order to analyze intrusion.
+Similar to Recovery phase approach taken in~\cite{zhang2009b}, Taser uses the Resolver model proposed to go back to the state of the operating system before the intrusion.
 
